---
title: "R Workshop"
output: 
  flexdashboard::flex_dashboard:
    theme: united
    css: style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

Introduction
==================================
Column {.tabset}
--------------------

### Introduction 


#### CS1B

The aim of Actuarial Statistics (CS1) is to provide a grounding in mathematical and statistical methods that are of relevance for actuarial work. It equips the student with knowledge of statistical distributions, methods to summarise data, the principles of statistical inference, regression models (including generalised linear models) and the fundamental concepts of Bayesian statistics.  The subject includes both theory and application of the ideas using R. 

#### CS2B

Risk Modelling and Survival Analysis (CS2) builds on CS1.  It develops knowledge of and the ability to apply statistical methods for risk modelling, time series analysis methods, stochastic processes (especially Markov chains and Markov jump processes), survival analysis (including regression methods applied to duration data), and graduation methods  It also includes a high-level introduction to machine learning. The subject includes both theory and application of the ideas to real data sets using R. 




```{r child="actuarial/CS1B-Paper-A.Rmd"}
```




```{r child="actuarial/CS1B-Paper-B.Rmd"}
```



```{r child="actuarial/CS1B-Paper-C.Rmd"}
```

```{r child="actuarial/CS1B-Paper-D.Rmd"}
```


```{r child="actuarial/CS1B-Paper-E.Rmd"}
```


```{r child="actuarial/CS1B-Paper-F.Rmd"}
```

```{r child="actuarial/CS1B-Paper-G.Rmd"}
```


```{r child="actuarial/CS1B-Paper-H.Rmd"}
```


```{r child="actuarial/CS1B-Paper-I.Rmd"}
```


```{r child="actuarial/CS1B-Paper-J.Rmd"}
```


```{r child="actuarial/CS2B-Paper-A.Rmd"}
```


```{r child="actuarial/CS2B-Paper-B.Rmd"}
```

```{r child="actuarial/CS2B-Paper-C.Rmd"}
```


```{r child="actuarial/CS2B-Paper-D.Rmd"}
```

```{r child="actuarial/CS2B-Paper-E.Rmd"}
```

```{r child="actuarial/CS2B-Paper-F.Rmd"}
```


```{r child="actuarial/CS2B-Paper-G.Rmd"}
```

```{r child="actuarial/CS2B-Paper-H.Rmd"}
```


```{r child="actuarial/CS2B-Paper-I.Rmd"}
```


```{r child="actuarial/CS2B-Paper-J.Rmd"}
```

