## Exercise 22}

1. The following are measurements (in mm) of a critical
dimension on a sample of engine crankshafts:

  \begin{array}{|cccc|}

    224.120 & 224.017 & 223.976 & 223.961 \\
    224.089 & 223.982 & 223.980 & 223.989  \\
    223.960 & 223.902  & 223.987 & 224.001  \\

  \end{array}

(a) Calculate the mean and standard deviation for these data.\\
(b) The process mean is supposed to be $\mu$ = 224mm. Is this the
case? Give reasons for your answer.
(c) Construct a 99\% confidence interval for these data and
interpret.
(d) Check that the normality assumption is valid using 2 suitable
plots.

<p>

%--------------------------------------------------------------%

* Exercise: Multiple plots in a single graphic window.
* Exercise: Adding Legends to plots.

* Exercise: Linear Regression models.
* Exercise: Correlation test.
* Exercise: Shapiro Wilk test.
* Exercise: Paired $t-$ test.


%--------------------------------------------------------------%

Forthcoming material

* plots (Continuing from before easter)

  * Par() command to set graphical parameters.
  * Box plots
\end{itemize}
* Linear Regression (lm()).
* Confidence intervals
* Hypothesis testing(p-values)
* Inference on Linear regression
* Special tests (Shapiro - Wilk test, correlation test)
* outliers
* Analysis of variance (ANOVA)





%--------------------------------------------------------------%

* Week 10  One class - Thursday 8th April
* Week 11  Two classes on Monday 12th April

* deadline for Assignment 4a Monday 8pm
* email kevin.obrien@ul.ie
\end{itemize}
* Week 11  One class on Thursday 15th April
* Week 12  Two classes on Monday 19th April
* Week 12  One class on Thursday 22nd April
* Week 13  Assignment 5 on Monday 26th April

* Short questions
* material from Week 11 onwards ( last 7 classes)
* open book exam
* worth 15%
\end{itemize}
* Week 13  one class on Monday 26th April
* Week 13  one class on Thursday 28th April
* Week 14  deadline for Assignment 6
\end{itemize}
%--------------------------------------------------------------%


\




1) Create a data frame for the data ( see `Car.R')
\begin{verbatim}
> Car<-data.frame(Make,Model,Cylinder,Weight,Mileage,Type)
\end{verbatim}

2) Find the `names' and the dimensions of the Car data frame.
\begin{verbatim}
> names(Car)
[1] "Make" "Model" "Cylinder" "Weight" "Mileage" "Type"
> dim(Car)
[1] 10  6
\end{verbatim}

3) Find the mean and standard deviation of the weight of the cars.
\begin{verbatim}
> mean(Car$Weight)
[1] 2858
> sd(Car$Weight)
[1] 543.7943
\end{verbatim}

4) Subset the data so that only cars above 2500 kgs are considered. (Call the subset 'cars2500')
\begin{verbatim}
cars2500 = Car[Car$Weight>2500,]
\end{verbatim}

5) Make a subset of the small cars (call the subset `smallcar').
\begin{verbatim}
> smallcar=Car[Car$Type=="Small",]
> smallcar
        Make  Model Cylinder Weight Mileage  Type
3       Ford Escort       V4   2345      33 Small
4      Eagle Summit       V4   2560      33 Small
5 Volkswagen  Jetta       V4   2330      26 Small
>
\end{verbatim}

6) What is the correlation coefficient between weight and mileage?

\begin{verbatim}
> cor(Car$Weight,Car$Mileage)
[1] -0.8759577
\end{verbatim}

7) Make a plot of weights against mileage (include a title).

\begin{verbatim}
>plot(Car$Weight, Car$Mileage, main="Weight vs. Mileage")
\end{verbatim}

Alternatively
\begin{verbatim}
>attach(Car)
>plot(Weight, Mileage, main="Weight vs. Mileage")
\end{verbatim}






\begin{verbatim}
# Car.R
> Make<-c("Honda","Chevrolet","Ford","Eagle","Volkswagen",
+"Buick","Mitsbusihi", "Dodge","Chrysler","Acura")
>
> Model<-c("Civic","Beretta","Escort","Summit","Jetta",
+"Le Sabre","Galant", "Grand Caravan","New Yorker","Legend")
>
> rep("V6",3)               # repeat three times
[1] "V6" "V6" "V6"
> Cylinder<-c(rep("V4",5),"V6","V4",rep("V6",3))
> Cylinder
[1] "V4" "V4" "V4" "V4" "V4" "V6" "V4" "V6" "V6" "V6"
> Weight<-c(2170,2655,2345,2560,2330,3325,2745,3735,3450,3265)
> Mileage<-c(33,26,33,33,26,23,25,18,22,20)
> Type<-c("Sporty","Compact",rep("Small",3),"Large",
+"Compact","Van",rep("Medium",2))
\end{verbatim}

<p>
