---
title: "``brewer_pal`` Colour Brewer palette (discrete)."
author: "Kevin O'Brien"
date: "19 October 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(scales)
```



#### Description
Colour Brewer palette (discrete).

#### Usage
<pre><code>
brewer_pal(type = "seq", palette = 1, direction = 1)
</code></pre>

#### Arguments
* ``type``: One of seq (sequential), div (diverging) or qual (qualitative)
* ``palette``: If a string, will use that named palette. If a number, will index into the list of palettes of appropriate type
* ``direction``: Sets the order of colours in the scale. If 1, the default, colours are as output by RColorBrewer::brewer.pal(). If -1, the order of colours is reversed.

#### References
http://colorbrewer2.org

#### Examples

```{r}
show_col(brewer_pal()(10))
show_col(brewer_pal("div")(5))
show_col(brewer_pal(palette = "Greens")(5))
# Can use with gradient_n to create a continous gradient
cols <- brewer_pal("div")(5)
show_col(gradient_n_pal(cols)(seq(0, 1, length.out = 30)))
```
