---
title: "``viridis_pal`` Viridis palette"
author: "Kevin O'Brien"
date: "19 October 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(scales)
```

#### Description
Viridis palette

#### Usage

<pre><code>
viridis_pal(alpha = 1, begin = 0, end = 1, direction = 1,
option = "D")
</code></pre>

#### Examples

```{r}
show_col(viridis_pal()(10))
show_col(viridis_pal(direction = -1)(6))
show_col(viridis_pal(begin = 0.2, end = 0.8)(4))
show_col(viridis_pal(option = "plasma")(6))
```

#### Arguments

* alpha The alpha transparency, a number in [0,1], see argument alpha in hsv.
* begin The (corrected) hue in [0,1] at which the viridis colormap begins.
* end The (corrected) hue in [0,1] at which the viridis colormap ends.
* direction Sets the order of colors in the scale. If 1, the default, colors are ordered from
darkest to lightest. If -1, the order of colors is reversed.
* option A character string indicating the colormap option to use. Four options are available:
"magma" (or "A"), "inferno" (or "B"), "plasma" (or "C"), "viridis" (or "D", the default option) and "cividis" (or "E").

#### References

* https://bids.github.io/colormap/

