``number_format`` Number formatters
#### Description
A set of functions to format numeric values:
*  number_format() and number() are generic formatters for numbers.
*  comma_format() and comma() format numbers with commas separating thousands.
*  percent_format() and percent() multiply values by one hundred and display percent sign.
*  unit_format() add units to the values.

All formatters allow you to re-scale (multiplicatively), to round to specified accuracy, to add
custom suffix and prefix and to specify decimal.mark and big.mark.

#### Usage
<pre><code>
number_format(accuracy = 1, scale = 1, prefix = "", suffix = "",
              big.mark = " ", decimal.mark = ".", trim = TRUE, ...)
number(x, accuracy = 1, scale = 1, prefix = "", suffix = "",
       big.mark = " ", decimal.mark = ".", trim = TRUE, ...)
comma_format(accuracy = 1, scale = 1, prefix = "", suffix = "",
             big.mark = ",", decimal.mark = ".", trim = TRUE, digits, ...)
comma(x, accuracy = 1, scale = 1, prefix = "", suffix = "",
      big.mark = ",", decimal.mark = ".", trim = TRUE, digits, ...)
percent_format(accuracy = NULL, scale = 100, prefix = "",
               suffix = "%", big.mark = " ", decimal.mark = ".", trim = TRUE,
               ...)
percent(x, accuracy = NULL, scale = 100, prefix = "",
        suffix = "%", big.mark = " ", decimal.mark = ".", trim = TRUE,
        ...)
unit_format(accuracy = 1, scale = 1, prefix = "", unit = "m",
            sep = " ", suffix = paste0(sep, unit), big.mark = " ",
            decimal.mark = ".", trim = TRUE, ...)
</code></pre>
#### Arguments
accuracy Number to round to, NULL for automatic guess.
scale A scaling factor: x will be multiply by scale before formating (useful if the underlying
                                                                      data is on another scale, e.g. for computing percentages or thousands).
28 number_format
prefix, suffix Symbols to display before and after value.
big.mark Character used between every 3 digits to separate thousands.
decimal.mark The character to be used to indicate the numeric decimal point.
trim Logical, if FALSE, values are right-justified to a common width (see base::format()).
... Other arguments passed on to base::format().
x A numeric vector to format.
digits Deprecated, use accuracy instead.
unit The units to append.
sep The separator between the number and the unit label.
Value
*_format() returns a function with single parameter x, a numeric vector, that returns a character
vector.
#### Examples
```{r}`
# number()
v <- c(12.3, 4, 12345.789, 0.0002)
number(v)
number(v, big.mark = ",")
number(v, accuracy = .001)
number(v, accuracy = .001, decimal.mark = ",")
number(v, accuracy = .5)
```
```{r}
# number_format()
my_format <- number_format(big.mark = "'", decimal.mark = ",")
my_format(v)
# comma() and comma_format()
comma_format()(c(1, 1e3, 2000, 1e6))
comma_format(accuracy = .01)(c(1, 1e3, 2000, 1e6))
comma(c(1, 1e3, 2000, 1e6))
```
```{r}
# percent() and percent_format()
percent_format()(runif(10))
percent(runif(10))
per_mille <- percent_format(
  scale = 1000,
  suffix = "\u2030",
  accuracy = .1
)
per_mille(.1234)
```
```{r}

french_percent <- percent_format(
  decimal.mark = ",",
  suffix = " %"
)
ordinal_format 29
french_percent(runif(10))
```
```{r}
# unit_format()
# labels in kilometer when the raw data are in meter
km <- unit_format(unit = "km", scale = 1e-3, digits = 2)
km(runif(10) * 1e3)
# labels in hectares, raw data in square meters
ha <- unit_format(unit = "ha", scale = 1e-4)
km(runif(10) * 1e5)
```
