---
title: "Forecasting with R"
subtitle: "Time Series Analysis with R"
author: DragonflyStats.github.io
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=3)
```


## Validating the Proportional Hazards Model

We have skipped much of the statistical theory up to this point, but we now
need to look at using some of it.

A common statistical approach for assessing regression models is through
analysis of the residuals, typically defined as some form of difference
between the predicted and observed values.

For survival analysis, our approach is more involved in terms of how we
define the residuals of a model, but the basic concept is the same.

### Fitting the Null Model

We first work from a null CoxPH model that we fit - the 'null' model:

```{r fit_null_model, echo=TRUE}
telco_model00_coxph <- coxph(Surv(accountdur, churned) ~ 1
                            ,data = telco_churn_tbl
                             )

telco_model00_coxph %>% summary()
```


### Martingale Residuals

We define our basic residual $m_i$ as follows:

$$
m_i = \delta_i - \hat{H}_0(t_i) \exp(x_i \hat{\beta})
$$

The residual is the difference between the observed value of the whether the
event occurred and its expected value under the Cox model.

```{r calculate_null_model_martingale_residuals, echo=TRUE}
telco_model00_resid_m <- telco_model00_coxph %>% residuals(type = 'martingale')

telco_model00_resid_m %>% summary()
```

We have zero-mean, but there is skew in the data as the values are bounded
above at 1, but unbounded below. The easiest way to check the shape is look at
a histogram of the values.

```{r plot_null_model_martingale_residuals, echo=TRUE}
ggplot() +
    geom_histogram(aes(x = telco_model00_resid_m), bins = 50) +
    xlab('Residual Value') +
    ylab('Count') +
    ggtitle('Histogram of Null-Model Martingale Residuals')

ggplot() +
    geom_histogram(aes(x = telco_model00_resid_m^2), bins = 50) +
    xlab('Squared Residual Value') +
    ylab('Count') +
    ggtitle('Histogram of Null-Model Squared Martingale Residuals')
```


### Deviance Residuals

Martingale residuals cannot be used in sums of squares methods to assess
goodness-of-fit - instead we define another quantity when the model is correct
is symmetrically distributed with an expectation of zero. These 'deviance'
residuals, $d_i$ are defined as follows:

$$
d_i = \text{sign}(m_i) {-2 [m_i + \delta_i \log (\delta_i - m_i)]}^{1/2}
$$

```{r calculate_null_model_deviance_residuals, echo=TRUE}
telco_model00_resid_d <- telco_model00_coxph %>% residuals(type = 'deviance')

telco_model00_resid_d %>% summary()
```

The mean is small, but non-zero. We inspect the distribution of values to see
what it looks like.


```{r plot_null_model_deviance_residuals, echo=TRUE}
ggplot() +
    geom_histogram(aes(x = telco_model00_resid_d), bins = 50) +
    xlab('Residual Value') +
    ylab('Count') +
    ggtitle('Histogram of Null-Model Deviance Residuals')

ggplot() +
    geom_histogram(aes(x = telco_model00_resid_d^2), bins = 50) +
    xlab('Squared Residual Value') +
    ylab('Count') +
    ggtitle('Histogram of Null-Model Squared Deviance Residuals')
```

Neither look distributed as we would like, suggesting this model is not good,
but we will return to this topic later.


### Residual Plots

Much like for linear models, these residuals are useful to help decide which
variables may have predictive power in our survival model. We do this by
plotting the values of the residual against the variable values.

To do this, we first add the residuals to the original dataset so that we can
construct some of the residual plots.

```{r create_residual_plots_null_model, echo=TRUE}
telco_null_assess_tbl <- telco_churn_tbl %>%
    mutate(resid_m = telco_model00_resid_m
          ,resid_d = telco_model00_resid_d
           )
```

Now that we have put the table together we can construct some two-way plots
of the variable against the martingale residual:


```{r construct_variable_residual_boxplots, echo=TRUE}
ggplot(telco_null_assess_tbl) +
    geom_boxplot(aes(x = vmailplan, y = resid_m)) +
    xlab('vmailplan') +
    ylab('Martingale Residual') +
    ggtitle('Residual Plot of Martingal Residual against vmailplan')

ggplot(telco_null_assess_tbl) +
    geom_boxplot(aes(x = intlplan, y = resid_m)) +
    xlab('intlplan') +
    ylab('Martingale Residual') +
    ggtitle('Residual Plot of Martingal Residual against intlplan')

ggplot(telco_null_assess_tbl) +
    geom_boxplot(aes(x = areacode, y = resid_m)) +
    xlab('areacode') +
    ylab('Martingale Residual') +
    ggtitle('Residual Plot of Martingal Residual against areacode')


ggplot(telco_null_assess_tbl) +
    geom_boxplot(aes(x = custservicecalls, y = resid_m, group = custservicecalls)) +
    xlab('custservicecalls') +
    ylab('Martingale Residual') +
    ggtitle('Residual Plot of Martingal Residual against custservicecalls')
```

`custservicecalls` is a numeric variable, but it does not take many values so
we can create boxplots rather than a more typical scatterplot.

For continuous variables, we look at those scatterplots:

```{r construct_variable_residual_scatterplots, echo=TRUE}
ggplot(telco_null_assess_tbl) +
    geom_point(aes(x = vmailmessage, y = resid_m), size = 1, alpha = 0.1) +
    xlab('vmailmessage') +
    ylab('Martingale Residual') +
    ggtitle('Residual Plot of Martingal Residual against vmailmessage')

ggplot(telco_null_assess_tbl) +
    geom_point(aes(x = daymins, y = resid_m), size = 1, alpha = 0.1) +
    xlab('daymins') +
    ylab('Martingale Residual') +
    ggtitle('Residual Plot of Martingal Residual against daymins')

ggplot(telco_null_assess_tbl) +
    geom_point(aes(x = nightmins, y = resid_m), size = 1, alpha = 0.1) +
    xlab('nightmins') +
    ylab('Martingale Residual') +
    ggtitle('Residual Plot of Martingal Residual against nightmins')

ggplot(telco_null_assess_tbl) +
    geom_point(aes(x = intlmins, y = resid_m), size = 1, alpha = 0.1) +
    xlab('intlmins') +
    ylab('Martingale Residual') +
    ggtitle('Residual Plot of Martingal Residual against intlmins')

```

No strong patterns stand out here, but that does not necessarily rule them
out of candidate models. We will not know the predictive power till we try
them in a model.


## Assessing the Proportional Hazards Assumption

The proportional hazards assumption is key in the construction of the partial
likelihood, as it allows the cancellation of the baseline hazard in the
calculations.

This assumption of proportional hazards is likely to be violated in practice,
so the question is how big the effect those violations have on our model.
Formal statistical tests are thus of limited value, but can be a useful
indicator of issues with any models built.


### Residual Plots

Tests for the proportional hazards assumption reduce to checking if the
parameters from the model are constant over time. This task is made easier
by the construction of 'Schoenfeld' residuals, which we will not discuss
here, but are easily calculated in R by using the `cox.zph()` function.

```{r testing_model01_ph_assumptions, echo=TRUE}
telco_model01_coxzph <- telco_model01_coxph %>% cox.zph(transform = 'km')

telco_model01_coxzph %>% print()
telco_model01_coxzph %>% ggcoxzph()
```

The plot and $p$-value suggest the assumptions are not horrible here.

We now move on to the second model we built.

```{r testing_model02_ph_assumptions, echo=TRUE}
telco_model02_coxzph <- telco_model02_coxph %>% cox.zph(transform = 'km')

telco_model02_coxzph %>% print()
telco_model02_coxzph %>% ggcoxzph()
```

The assumption also seems sound here.


We now try the combined model, `model03`:

```{r testing_model03_ph_assumptions, echo=TRUE}
telco_model03_coxzph <- telco_model03_coxph %>% cox.zph(transform = 'km')

telco_model03_coxzph %>% print()
telco_model03_coxzph %>% ggcoxzph()
```



# R Environment

```{r show_session_info, echo=TRUE, message=TRUE}
sessioninfo::session_info()
```