---
title: "The Survival Functions"
subtitle: "Survival Analysis with R"
author: DragonflyStats.github.io
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=3)
```




#### Smoker Data

The smoker data is a smoking relapse study involving a number of different
treatments and details on when and if the patient resumed smoking.

```{r show_smoker_data, echo=TRUE}
pharmaco_smoker_tbl %>% glimpse()

pharmaco_smoker_tbl %>% head()
```


### Data Setup Script
library(tidyverse)
library(asaur)

# Re-using some code from Sec 1 of the workshop to avoid repeating myself
# in later worksheets.


# Setup the Telco churn data
telco_cols <- cols(
    areacode = col_character()
)

telco_churn_tbl <- read_csv('data/telcochurn.csv'
                           ,col_types = telco_cols
                            )

# Setup the pharmacoSmoking data
pharmaco_smoker_tbl <- asaur::pharmacoSmoking %>% as_data_frame()

# Setup the prostate survival data
prostate_surv_tbl <- asaur::prostateSurvival %>% as_data_frame()