---
title: "Forecasting with R"
subtitle: "Time Series Analysis with R"
author: DragonflyStats.github.io
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=3)
```

# Comparing Curves: The Log-Rank Test

Now that we have ways of constructing different survival curves from our data
we turn our attention to checking if these differences can be explained by
random chance


|            |Control          |    Treatment      |   Total     |
|----------- |:---------------:|:-----------------:|:-----------:|
|Failure     |$d_{0i}$         | $d_{1i}$          | $d_i$       |
|Non-failure |$n_{0i} - d_{0i}$| $n_{1i} - d_{1i}$ | $n_i - d_i$ |
|Total       |$n_{0i}$         | $n_{1i}$          | $n_i$       |


From this table of values, we define two values:

\begin{eqnarray*}
e_{0i} &=& E(d_{0i}) = \frac{n_{0i} d_i}{n_i}                                        \\
v_{0i} &=& \text{var}(d_{0i}) = \frac{n_{0i} n_{1i} d_i (n_i - d_i)}{n_i^2(n_i - 1)} \\ 
U_0    &=& \sum_{i=1}^{N} (d_{0i} - e_{0i}) = \sum d_{0i} - \sum e_{0i}              \\
V_0    &=& \text{var}(U_0) = \sum v_{0i}
\end{eqnarray*}

Having defined these variables as above, we can show that

$$
\frac{U_0^2}{V_0} \sim \chi_1^2
$$

This gives us a natural statistical test to compare the two curves.


## Comparing Binary Categories

The simplest case is when we are comparing survival data split down binary
outcomes, such as whether or not the customer has a voicemail plan or a plan
with international calls.

### Voicemail Differences

We start by comparing the voicemail plans.

```{r compare_vmail_curves, echo=TRUE}
telco_vmail_survdiff <- survdiff(Surv(accountdur, churned) ~ vmailplan
                                ,data = telco_churn_tbl)

telco_vmail_survdiff %>% print()
telco_vmail_survdiff %>% glance()
```

Such a small $p$-value suggests the differences in the curves are unlikely to
be due to chance, and this gels with what we observed on the plots of the
curves.

### International Calls Differences

Now we compare the differences between plans with and without international
calls.

```{r compare_intl_curves, echo=TRUE}
telco_intl_survdiff <- survdiff(Surv(accountdur, churned) ~ intlplan
                               ,data = telco_churn_tbl)

telco_intl_survdiff %>% print()
telco_intl_survdiff %>% glance()
```

## Comparing Multiple Categories

The approach discussed can be extended to categorical variables with multiple
values. We can use `areacode` for this as it has three values in it:

```{r check_area_code_levels, echo=TRUE}
telco_churn_tbl %>% count(areacode)
```

Let's check how it splits the survival curves.

```{r create_areacode_km_estimates, echo=TRUE}
telco_areacode_km <- survfit(Surv(accountdur, churned) ~ areacode
                            ,data = telco_churn_tbl)

telco_areacode_km %>% ggsurvplot(censor = TRUE, conf.int = TRUE)
```

Now let's run the log-rank test on this data:

```{r compare_areacode_curves, echo=TRUE}
telco_areacode_survdiff <- survdiff(Surv(accountdur, churned) ~ areacode
                                   ,data = telco_churn_tbl)

telco_areacode_survdiff %>% print()
telco_areacode_survdiff %>% glance()
```

Again, the results of the log-rank test meshes with our visual inspection: a
$p$-value of 0.6 suggests differences are likely due to chance, and for most
of the curves the three area codes pretty much share the same experience.

Using our domain knowledge, this also makes sense - we would not expect phones
based in different area codes to have substantial differences: area codes are
too big and broad to show systemic geographic effects.


## Discretizing Continuous Variables

In the event that we have numeric variables affecting survival, we need to
perform some discretization of those values to allow us to assess these
effects.

Regression models discussed later will allow us tackle these dependencies in
a more direct way, but for now we will use this simplified approach.

One numeric variable likely to have an effect on churn is `custservicecalls` -
the number of calls made to customer service. We would expect that an account
with more service calls is likely to be dissatisfied, so higher counts of
calls corresponds to a higher churn rate.

### Discretizing `custservicecalls`

Having decided to bin our calls, we now need to decide how we go about this.
One way is to split the data by quantiles, so we can try this first.

Should this not prove satisfactory, it may help inform more suitable ways
to bin this value.

```{r create_custservicecalls_bins, echo=TRUE}
telco_churn_tbl %>% count(custservicecalls) %>% mutate(cuml = cumsum(n))
```

We see this is a very skewed distribution to the lower end of call counts, so
`ntile()` is not as effective here as it will not properly separate the
counts. Instead we use `cut()` as that gives us more control over the binning.


```{r construct_callservice_cuts, echo=TRUE}
telco_churn_cat_tbl <- telco_churn_tbl %>%
    mutate(calls_cat = cut(custservicecalls
                          ,breaks = c(0, 1, 2, 5, Inf)
                          ,right = FALSE)
    )

telco_callscat_km <- survfit(Surv(accountdur, churned) ~ calls_cat
                            ,data = telco_churn_cat_tbl)

telco_callscat_km %>% ggsurvplot(censor = FALSE, conf.int = TRUE)
```

It looks like there is not much difference for lower counts, but there is a
worse churn experience when the number of customer service calls is 5 or
higher.

```{r compare_callscat_curves, echo=TRUE}
telco_callscat_survdiff <- survdiff(Surv(accountdur, churned) ~ calls_cat
                                   ,data = telco_churn_cat_tbl)

telco_callscat_survdiff %>% print()
telco_callscat_survdiff %>% glance()
```


