
# Estimations of the Survival Functions

## The Kaplan-Meier Estimator

Kaplan-Meier is the standard method for estimating the survival function of a
given dataset. Formally, it is defined as follows

$$
\hat{S}(t) = \prod_{t_i \leq t} (1 - \hat{q}_i) = \prod_{t_i \leq t} \left(1 - \frac{d_i}{n_i}\right)
$$

where $n_i$ is the number of subjects at risk at time $t$, and $d_i$ is the
number of individuals who fail at that time.

### Using Kaplan-Meier

In R, we construct KM estimators using the `survfit()` function.

Before we move on to our datasets, we start with a small set of data.

```{r km_estimator_intro, echo=TRUE}
tt   <- c(7, 6, 6, 5, 2, 4)
cens <- c(0, 1, 0, 0, 1, 1)
grp  <- c(0, 0, 0, 1, 1, 1)

Surv(tt, cens)

sample_tbl <- tibble(tt = tt, cens = cens, grp = grp)

example_km <- survfit(Surv(tt, cens) ~ 1, data = sample_tbl, conf.type = 'log-log')

plot(example_km)
```

Basic plotting routines are worth trying, but the `survminer` package has
specialised plots that use `ggplot2` to create them.

```{r km_estimator_survminer, echo=TRUE}
ggsurvplot(example_km)
```

Printing out the 'fitted' object gives us some basic statistics:

```{r km_estimator_print, echo=TRUE}
example_km %>% print()
```

We get more details from the `summary()` function:

```{r km_estimator_summary, echo=TRUE}
example_km %>% summary()
```

#### Exercises

  1. Construct the KM estimator for the telco churn data
  1. What is the median survival time for this data?
  1. What is the mean survival time?
  1. Repeat the above for the other two datasets.


### Follow-up Time

A useful measure may be how long the observation period lasts, something that
can be subtly difficult to measure.

One method is to switch the censoring labels - that is, we consider the
original event as a censoring of the observation in the study.

```{r measure_followup_time, echo=TRUE}
sample_tbl <- sample_tbl %>%
    mutate(follow = 1 - cens)

follow_km <- survfit(Surv(tt, follow) ~ 1, data = sample_tbl, conf.type = 'log-log')

ggsurvplot(follow_km)

follow_km %>% summary()
```


## Smoothed Hazard Functions

An empirical estimate of the hazard function is given by

$$
\mu(i) = \frac{d_i}{n_i}
$$

This is a very noisy estimate as it is sensitive to sample noise.

### The Hazard Estimator

To obtain more smooth functions we use kernel density estimator techniques.

```{r show_muhaz_plots, echo=TRUE}
sample_muhaz <- muhaz(sample_tbl$tt, sample_tbl$cens, max.time = 7)

plot(sample_muhaz)
```

`broom` has tidying methods for `muhaz()` and this allows us to create plots
with `ggplot2`

```{r show_muhaz_tidy, echo=TRUE}
muhaz_tidy_tbl <- sample_muhaz %>% tidy()

muhaz_tidy_tbl %>% glimpse()
```

We have estimates of the hazard function now and so can plot it.

```{r show_muhaz_tidy_plot, echo=TRUE}
ggplot(muhaz_tidy_tbl) +
    geom_line(aes(x = time, y = estimate)) +
    expand_limits(x = 0, y = 0) +
    xlab("Time") +
    ylab("Estimated Hazard")
```


#### Exercises

  1. Construct the smoothed estimator for the telco churn data
  1. What time has the highest hazard rate?
  1. What time has the lowest hazard rate?
  1. Repeat the above for the other two datasets.


### Boundary Corrections

By default, `muhaz()` corrects for the boundary on both sides, but we may not
wish this. To get estimates without this correction, we add it as an argument
to the function call.


```{r show_muhaz_nocorr, echo=TRUE}
sample_nocorr_muhaz <- muhaz(sample_tbl$tt, sample_tbl$cens, max.time = 7
                            ,b.cor = 'none')

plot(sample_nocorr_muhaz)


sample_nocorr_muhaz_tidy_tbl <- sample_nocorr_muhaz %>% tidy()

ggplot(sample_nocorr_muhaz_tidy_tbl) +
    geom_line(aes(x = time, y = estimate)) +
    expand_limits(x = 0, y = 0) +
    xlab("Time") +
    ylab("Estimated Hazard")
```

To help ensure these smoothed estimates are capturing the correct aspects of
the data, we also have the equivalent `pehaz()` functions - giving histogram
estimates of the hazards, much how histogram and kernel estimates are discrete
and continuous analogies of one another.

#### Exercises

  1. Construct the non-corrected smoothed estimator for the telco churn data
  1. Repeat the above for the other two datasets.


## Comparing KM and Smoothed Estimates

Now that we have both methods of checking the empirical estimates, it is good
to compare them.

Before we do this, we note that Kaplan-Meier estimates the survival function
but `muhaz()` gives us estimates of the hazard function.

Thus we need to do some conversions and the easiest way to do this is to
numerically integrate the hazard functions to calculate the survival function
and then compare to the KM estimate.

```{r integrate_muhaz, echo=TRUE}
muhaz_surv_tbl <- muhaz_tidy_tbl %>%
    mutate(dt  = c(0, diff(time))
          ,S_t = exp(-cumsum(estimate * dt))
    )

ggsurvplot(example_km)$plot +
    geom_line(aes(x = time, y = S_t), data = muhaz_surv_tbl)
```


### No Boundary Correction

We also want to compare the results without the boundary correction.

```{r compare_nocorr_estimates, echo=TRUE}
muhaz_nocorr_surv_tbl <- sample_nocorr_muhaz_tidy_tbl %>%
    mutate(dt  = c(0, diff(time))
          ,S_t = exp(-cumsum(estimate * dt))
    )

ggsurvplot(example_km)$plot +
    geom_line(aes(x = time, y = S_t), colour = 'blue', data = muhaz_surv_tbl) +
    geom_line(aes(x = time, y = S_t), data = muhaz_nocorr_surv_tbl)
    

```

We see that removing the boundary corrections can cause big discrepancies
between the smoothed and discrete estimates (at least in this case)


## Estimating Survival Curves for the Telco Churn Data

We now are going to build various Kaplan-Meier estimates for all the data,
as well as splits on the data along `vmailplan` - whether or not the account
had a voice mail, and `intlplan` - whether or not the account had an
international calls plan.

```{r create_km_fits, echo=TRUE}
telco_all_km   <- survfit(Surv(accountdur, churned) ~ 1,         data = telco_churn_tbl)
telco_vmail_km <- survfit(Surv(accountdur, churned) ~ vmailplan, data = telco_churn_tbl)
telco_intl_km  <- survfit(Surv(accountdur, churned) ~ intlplan,  data = telco_churn_tbl)
```


### Plotting the Curves

Having created the KM estimates, we first plot the curves to check for
differences.

```{r plot_km_curves, echo=TRUE}
telco_all_km   %>% ggsurvplot(censor = FALSE)
telco_vmail_km %>% ggsurvplot(censor = FALSE)
telco_intl_km  %>% ggsurvplot(censor = FALSE)
```

The estimators comes with upper and lower confidence bounds, and it may be
useful to include these estimates when determining the differences. We will
look at a more formal way of discriminating between curves in a while.

```{r plot_km_curves_bounds, echo=TRUE}
telco_all_km   %>% ggsurvplot(censor = FALSE, conf.int = TRUE)
telco_vmail_km %>% ggsurvplot(censor = FALSE, conf.int = TRUE)
telco_intl_km  %>% ggsurvplot(censor = FALSE, conf.int = TRUE)
```


### Combining KM Estimators

This is useful, but it would be better to combine all these together. To do
this, we need the tables of the estimator data, and stack them together.

Fortunately, the *broom* package is compatible with basic survival analysis
routines, so this provides us with some utilities to help with this.

If we need to access the KM data directly, we can use `tidy()` to get a look:

```{r introspect_km_fits, echo=TRUE}
telco_all_km_data   <- telco_all_km   %>% tidy()
telco_all_km_data   %>% glimpse()

telco_vmail_km_data <- telco_vmail_km %>% tidy()
telco_vmail_km_data %>% glimpse()

telco_intl_km_data  <- telco_intl_km  %>% tidy()
telco_intl_km_data  %>% glimpse()
```

We stack the data and create a line plot of each of the survival curves.

```{r plot_stacked_km_plots, echo=TRUE}
stacked_km_tbl <- list(telco_all_km_data %>% mutate(strata = 'ALL')
                      ,telco_vmail_km_data
                      ,telco_intl_km_data
                       ) %>%
    bind_rows()


ggplot(stacked_km_tbl) +
    geom_line(aes(x = time, y = estimate,  colour = strata)) +
    geom_line(aes(x = time, y = conf.high, colour = strata), linetype = 'dashed') +
    geom_line(aes(x = time, y = conf.low,  colour = strata), linetype = 'dashed')
```

This chart is not as useful as it could be - it is too 'busy'. We need to
simplify.

Instead of including everything, we will look at the international plan curves
along with the estimator calculated from the full dataset.

```{r plot_stacked_km_all_intl_plots, echo=TRUE}
ggplot(stacked_km_tbl %>% filter(grepl('ALL|intl', strata))) +
    geom_line(aes(x = time, y = estimate,  colour = strata)) +
    geom_line(aes(x = time, y = conf.high, colour = strata), linetype = 'dashed') +
    geom_line(aes(x = time, y = conf.low,  colour = strata), linetype = 'dashed')
```

From this, it looks like people without international plans are similar to
the population as a whole, but those with plans have a much worse churn
effect. This type of information will be useful later.

We can do a similar thing for the voice-mail plans.

```{r plot_stacked_km_all_vmail_plots, echo=TRUE}
ggplot(stacked_km_tbl %>% filter(grepl('ALL|vmail', strata))) +
    geom_line(aes(x = time, y = estimate,  colour = strata)) +
    geom_line(aes(x = time, y = conf.high, colour = strata), linetype = 'dashed') +
    geom_line(aes(x = time, y = conf.low,  colour = strata), linetype = 'dashed')
```

This split is more mixed, but it appears that those with a voicemail plan have
a better churn experience than the population, but those without are similar
to the population as a whole.




