---
title: "cdata"
author: "Kevin O'Brien"
date: "28 October 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(cdata)
```

## ``blocks_to_rowrecs_q``: Map sets rows to columns (query based, take name of table).

#### Description 

Transform data facts from rows into additional columns using SQL and controlTable.

#### Usage 
<pre><code>
blocks_to_rowrecs_q(tallTable, keyColumns, controlTable, my_db, ...,
columnsToCopy = NULL,
tempNameGenerator = mk_tmp_name_source("mvtcq"), strict = FALSE,
checkNames = TRUE, showQuery = FALSE, defaultValue = NULL,
dropDups = FALSE, temporary = FALSE, resultName = NULL)
</code></pre>

#### Arguments 
* tallTable name of table containing data to be mapped (db/Spark data)
* keyColumns character list of column defining row groups
* controlTable table specifying mapping (local data frame)
* my_db db handle
* ... force later arguments to be by name.
* ``columnsToCopy``: character list of column names to copy
* tempNameGenerator: a tempNameGenerator from cdata::mk_tmp_name_source()
* strict logical, if TRUE check control table name forms
* checkNames logical, if TRUE check names
* showQuery if TRUE print query
* defaultValue if not NULL literal to use for non-match values.
* dropDups logical if TRUE suppress duplicate columns (duplicate determined by name, not
content).
* temporary logical, if TRUE make result temporary.
* resultName character, name for result table.

#### Details 

This is using the theory of "fluid data"n (https://github.com/WinVector/cdata), which in-
cludes the principle that each data cell has coordinates independent of the storage details 
 and storage detail dependent coordinates (usually row-id, column-id, and group-id) can be re-derived at will
(the other principle is that there may not be "one true preferred data shape" and many re-shapings
of data may be needed to match data to different algorithms and methods).
The controlTable defines the names of each data element in the two notations: the notation of the
tall table (which is row oriented) and the notation of the wide table (which is column oriented).
controlTable[ , 1] (the group label) cross colnames(controlTable) (the column labels) are names of
data cells in the long form. controlTable[ , 2:ncol(controlTable)] (column labels) are names of data
cells in the wide form. To get behavior similar to tidyr::gather/spread one builds the control table
by running an appropriate query over the data.



#### Examples 
```{r}
if (requireNamespace("DBI", quietly = TRUE) &&
requireNamespace("RSQLite", quietly = TRUE)) {
my_db <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")
# pivot example
d <- data.frame(meas = c('AUC', 'R2'), val = c(0.6, 0.2))
DBI::dbWriteTable(my_db,
'd',
d,
temporary = TRUE)
cT <- build_pivot_control_q('d',
columnToTakeKeysFrom= 'meas',
columnToTakeValuesFrom= 'val',
my_db = my_db)
tab <- blocks_to_rowrecs_q('d',
keyColumns = NULL,
controlTable = cT,
my_db = my_db)
qlook(my_db, tab)
DBI::dbDisconnect(my_db)
}
```

Some discussion and examples   can be found here: https://winvector.github.io/FluidData/
FluidData.html and here https://github.com/WinVector/cdata.
Value
wide table built by mapping key-grouped tallTable rows to one row per group6
build_pivot_control
See Also
rowrecs_to_blocks_q, build_pivot_control_q, blocks_to_rowrecs
