## ``build_pivot_control``

Build a ``blocks_to_rowrecs()``/``rowrecs_to_blocks()`` control table that specifies a pivot from a data.frame.

#### Description 


Some discussion and examples   can be found here: https://winvector.github.io/FluidData/
FluidData.html.

#### Usage 
 <pre><code>
build_pivot_control(table, columnToTakeKeysFrom, columnToTakeValuesFrom,
..., prefix = columnToTakeKeysFrom, sep = NULL)

## Default S3 method:
build_pivot_control(table, columnToTakeKeysFrom,
columnToTakeValuesFrom, ..., prefix = columnToTakeKeysFrom,
sep = NULL)

## S3 method for class 'relop'
build_pivot_control(table, columnToTakeKeysFrom,
columnToTakeValuesFrom, ..., prefix = columnToTakeKeysFrom,
sep = NULL, tmp_name_source = wrapr::mk_tmp_name_source("bpc"),
temporary = FALSE)
</code></pre>

#### Arguments 

* ``table``: data.frame to scan for new column names (in-memory data.frame).
* ``columnToTakeKeysFrom``: character name of column build new column names from.
* ``columnToTakeValuesFrom``: character name of column to get values from.
* ``...``: not used, force later args to be by name
* ``prefix``: column name prefix (only used when sep is not NULL)
* ``sep``: separator to build complex column names.
* ``tmp_name_source`: a tempNameGenerator from ``cdata::mk_tmp_name_source()``
* ``temporary``: logical, if TRUE use temporary tables



#### Examples 
```{r}
d <- data.frame(measType = c("wt", "ht"),
measValue = c(150, 6),
stringsAsFactors = FALSE)
build_pivot_control(d,
'measType', 'measValue',
sep = '_')
```

```{r}
d <- data.frame(measType = c("wt", "ht"),
measValue = c(150, 6),
stringsAsFactors = FALSE)
ops <- rquery::local_td(d) %.>%
build_pivot_control(.,
'measType', 'measValue',
sep = '_')
cat(format(ops))
if(requireNamespace("rqdatatable", quietly = TRUE)) {8
build_pivot_control_q
}
```

```{r}
library("rqdatatable")
d %.>%
ops %.>%
print(.)
if(requireNamespace("RSQLite", quietly = TRUE)) {
db <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")
DBI::dbWriteTable(db,
'd',
d,
overwrite = TRUE,
temporary = TRUE)
db %.>%
ops %.>%
print(.)
DBI::dbDisconnect(db)
}
```
#### Value
control table
See Also
blocks_to_rowrecs, build_pivot_control_q
