---
title: "cdata"
author: "Kevin O'Brien"
date: "28 October 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(cdata)
```

## ``build_pivot_control_q`` 

Build a ``blocks_to_rowrecs_q()`` control table that specifies a pivot (query based, takes name of table).

#### Description 


Some discussion and examples can be found here: https://winvector.github.io/FluidData/FluidData.html.

#### Usage 

<pre><code>
build_pivot_control_q(tableName, columnToTakeKeysFrom,
columnToTakeValuesFrom, my_db, ..., prefix = columnToTakeKeysFrom,
sep = NULL)
</code></pre>

#### Arguments 

* ``tableName``: Name of table to scan for new column names.
* ``columnToTakeKeysFrom``: character name of column build new column names from.
* ``columnToTakeValuesFrom``: character name of column to get values from.
* ``my_db``: db handle
* ``...``: not used, force later args to be by name
* ``prefix``: column name prefix (only used when sep is not NULL)
* ``sep``: separator to build complex column names.


#### Examples 
```{r}
if (requireNamespace("DBI", quietly = TRUE) &&
requireNamespace("RSQLite", quietly = TRUE)) {
my_db <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")
d <- data.frame(measType = c("wt", "ht"),
measValue = c(150, 6),
stringsAsFactors = FALSE)
DBI::dbWriteTable(my_db,
'd',
d,
overwrite = TRUE,
temporary = TRUE)
build_pivot_control_q('d', 'measType', 'measValue',
my_db = my_db,
sep = '_') %.>%
print(.)
DBI::dbDisconnect(my_db)
}
```


#### Value

control table

#### See Also

blocks_to_rowrecs_q, build_pivot_control_q

