---
title: "cdata"
author: "Kevin O'Brien"
date: "28 October 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(cdata)
```

map_fields

Map field values from one column into new derived columns (takes a
data.frame).

#### Description 

Map field values from one column into new derived columns (takes a data.frame).

#### Usage 

<pre><code>
map_fields(d, cname, m)
</code></pre>

#### Arguments 

* d name of table to re-map.
* cname name of column to re-map.
* m name of table of data describing the mapping (cname column is source, derived columns are destinations).



#### Examples 
```{r}
d <- data.frame(what = c("acc", "loss",
"val_acc", "val_loss"),
score = c(0.8, 1.2,
0.7, 1.7),
stringsAsFactors = FALSE)
m <- data.frame(what = c("acc", "loss",
"val_acc", "val_loss"),
measure = c("accuracy", "log-loss",
"accuracy", "log-loss"),
dataset = c("train", "train", "validation", "validation"),
stringsAsFactors = FALSE)
map_fields(d, 'what', m)
```

#### Value
re-mapped table12
map_fields_q
