---
title: "``pivot_to_rowrecs``: Move values from rows to columns (pivot)."
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(cdata)
```


#### Description 

This is a convenience notation for ``blocks_to_rowrecs``. 


#### Usage 
 <pre><code>
pivot_to_rowrecs(0data, columnToTakeKeysFrom, columnToTakeValuesFrom,
rowKeyColumns, ..., sep = NULL, checkNames = TRUE,
checkKeys = TRUE, strict = FALSE, use_data_table = FALSE)
</code></pre>

#### Arguments 

* ``data``: data.frame to work with (must be local, for remote please try moveValuesToColumns*).
* ``columnToTakeKeysFrom``: character name of column build new column names from.
* ``columnToTakeValuesFrom``: character name of column to get values from.
* ``rowKeyColumns``: character array names columns that should be table keys.
* ``...``: force later arguments  to bind by name.
* ``sep``: character if not null build more detailed column names.
* ``checkNames``: logical, if TRUE check names.
* ``checkKeys``: logical, if TRUE check keyColumns uniquely identify blocks (required).
* ``strict``: logical, if TRUE check control table name forms
* ``use_data_table``: logical if TRUE try to use data.table for the pivots.


#### Examples 
```{r}
d <- data.frame(meas= c('AUC', 'R2'), val= c(0.6, 0.2))
pivot_to_rowrecs(d,
columnToTakeKeysFrom= 'meas',
columnToTakeValuesFrom= 'val',
rowKeyColumns= c()) %.>%
print(.)
```

#### Value
new data.frame with values moved to columns.

#### See Also
unpivot_to_blocks, blocks_to_rowrecs
