---
title: "cdata"
author: "Kevin O'Brien"
date: "28 October 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(cdata)
```

## ``rowrecs_to_blocks``
Map a set of columns to rows (takes a data.frame).

#### Description 

Transform data facts from columns into additional rows controlTable.

#### Usage 
 <pre><code>
rowrecs_to_blocks(wideTable, controlTable, ..., checkNames = TRUE,
  checkKeys = FALSE, strict = FALSE, columnsToCopy = NULL,
  use_data_table = FALSE)

## Default S3 method:
rowrecs_to_blocks(wideTable, controlTable, ...,
  checkNames = TRUE, checkKeys = FALSE, strict = FALSE,
  columnsToCopy = NULL, use_data_table = FALSE)

## S3 method for class 'relop'
rowrecs_to_blocks(wideTable, controlTable, ...,
  checkNames = TRUE, strict = FALSE, columnsToCopy = NULL,
  use_data_table = TRUE,
  tmp_name_source = wrapr::mk_tmp_name_source("rrtbl"),
  temporary = TRUE)
</code></pre>

#### Arguments 
* ``wideTable`` data.frame containing data to be mapped (in-memory data.frame).
* ``controlTable``: table specifying mapping (local data frame).
* ``...``: force later arguments  to be by name.
* ``checkNames``: logical, if TRUE check names.
* ``checkKeys: logical, if TRUE check columnsToCopy form row keys (not a requirement, unless you want to be able to invert the operation).
* ``strict``: logical, if TRUE check control table name forms.
* ``columnsToCopy``: character array of column names to copy.
* ``use_data_table``: logical if TRUE try to use data.table for the un-pivots.
* ``tmp_name_source``: a tempNameGenerator from cdata::mk_tmp_name_source()
* ``temporary``: logical, if TRUE make result temporary.





#### Examples 
```{r}
# un-pivot example
d <- data.frame(AUC = 0.6, R2 = 0.2)
cT <- build_unpivot_control(nameForNewKeyColumn= 'meas',
  nameForNewValueColumn= 'val',
  columnsToTakeFrom= c('AUC', 'R2'))

rowrecs_to_blocks(d, cT)
```

```{r}
d <- data.frame(AUC = 0.6, R2 = 0.2)
cT <- build_unpivot_control(
  nameForNewKeyColumn= 'meas',
  nameForNewValueColumn= 'val',
  columnsToTakeFrom= c('AUC', 'R2'))

ops <- rquery::local_td(d) %.>%
rowrecs_to_blocks(., cT)
cat(format(ops))
```

```{r}
if(requireNamespace("rqdatatable", quietly = TRUE)) {
library("rqdatatable")

 d %.>%
 ops %.>%
 print(.)
 }
if(requireNamespace("RSQLite", quietly = TRUE)) {
db <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")
DBI::dbWriteTable(db,
 'd',
 d,
 overwrite = TRUE,
 temporary = TRUE)

 db %.>%
 ops %.>%
 print(.)
 DBI::dbDisconnect(db)
}
```

#### Details 

* This is using the theory of "fluid data" (https://github.com/WinVector/cdata), which includes
the principle that each data cell has coordinates independent of the storage details 
 and storage detail dependent coordinates (usually row-id, column-id, and group-id) can be re-derived at will (the other
principle is that there may not be "one true preferred data shape" and many re-shapings of data may
be needed to match data to different algorithms and methods).
* The ``controlTable`` defines the names of each data element in the two notations: the notation of the
tall table (which is row oriented) and the notation of the wide table (which is column oriented).
* controlTable[ , 1] (the group label) cross colnames(controlTable) (the column labels) are names of
data cells in the long form. 
* ``controlTable[ , 2:ncol(controlTable)]`` (column labels) are names of data
cells in the wide form. To get behavior similar to tidyr::gather/spread one builds the control table
by running an appropriate query over the data.

* Some discussion and examples can be found here: https://winvector.github.io/FluidData/
FluidData.html and here https://github.com/WinVector/cdata.


#### Value
long table built by mapping wideTable to one row per group
#### See Also
build_unpivot_control, blocks_to_rowrecs_q
