---
title: "MCMC with R"
author: "Kevin O Brien"
date: "18 December 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 3 Monte Carlo Integration

3.1 Introduction
3.2 Classical Monte Carlo Integration
3.3 Importance Sampling

## 3.1 Introduction


#### Page 63 and 64


As a test, we compare the use of ``integrate()`` on the integral
\[  \int^{\infty}_{0} x^{\lambda-1} exp(-x) dx   \]

with the computation of $\Gamma(\lambda)$ via the ``gamma()`` function.

```{r}
cac = rcauchy(10)
nin = function(a){ integrate(lik,-a,a)$val }
nan = function(a){ area(lik,-a,a)$val }
x = seq(1,10^3,10^2)
y = log(apply(as.matrix(x),1,nin))
z = log(apply(as.matrix(x),1,nan))
```


```{r}
plot(x,y,type="l", ylim = range(cbind(y,z)), lwd = 2)
lines(x,z,lty=2,col="sienna",lwd=2)
```

## 3.3 Importance Sampling

Importance sampling is a general technique for estimating properties of a particular distribution, while only having samples generated from a different distribution than the distribution of interest. It is related to umbrella sampling in computational physics. 

Depending on the application, the term may refer to the process of sampling from this alternative distribution, the process of inference, or both.

# Example 3.5

\[g(y) =  \frac{e^{-y} }{  \int^{\infty}_{4.5} e^{-x}dx}\]

#### Page 71

```{r}
Nsim = 1000
y = rexp(Nsim) + 4.5
weit = dnorm(y) / dexp(y - 4.5)
plot(cumsum (weit)/1:Nsim,type="l")
abline(a = pnorm(-4.5),b=0,col="red")
```

#### Page 72

```{r}

f = function(a,b){
 exp( 2*(lgamma(a+b)-lgamma(a)-lgamma(b)) + a*log(0.3) + b*log(0.2)) 
}

aa = 1:150
bb = 1:100

post = outer(aa,bb,f)
image(aa,bb,post,xlab=expression(alpha),ylab="")
contour(aa,bb,post=add=T)
```

#### Page 73

```{r}
# T sample
x = matrix( rt (2 * 10^4,3),ncol=2)

#Scale Matrix
E = matrix( c(220,190,190,180),ncol=2)

image(aa,bb,post,xlab=expression(alpha),ylab="")
y = t(t(chol(E) %*% t(x)) + c(50,45))
points(y,cex =0.6,pch=19)

```


#### Page 75


```{r}
ine = apply(y,1, min)

y = y[ine >0, ]
x = x[ine >0, ]

normx = sqrt(x[,1]^2 + x[,2]^2)

f = function(a) exp(2*(lgamma(a[,1] + a[,2]) - lgamma(a[,1]) - lgamma(a[,2]) + a[,1]*log(0.3) + a[,2]*log(0.2) )    

h = function(a) exp(1*(lgamma(a[,1] + a[,2]) - lgamma(a[,1]) - lgamma(a[,2]) + a[,1]*log(0.5) + a[,2]*log(0.5) )    
den = dt(normx,3)

mean( f(y) / den) / mean( h(y) / den) 

```