## Clean data.frame names with ``clean_names()``

Call this function every time you read data.

It works in a %>% pipeline, and handles problematic variable names, especially those that are so well-preserved 
by ``readxl::read_excel()`` and ``readr::read_csv()``.

*    Parses letter cases and separators to a consistent format.
*    .... Default is to snake_case, but other cases like camelCase are available
*    Handles special characters and spaces, including transliterating characters like œ to oe.
*    Appends numbers to duplicated names
*    Converts "%" to "percent" and "#" to "number" to retain meaning
*    Spacing (or lack thereof) around numbers is preserved

```{r}
# Create a data.frame with dirty names
test_df <- as.data.frame(matrix(ncol = 6))
names(test_df) <- c("firstName", "ábc@!*", "% successful (2009)",
                    "REPEAT VALUE", "REPEAT VALUE", "")
```
Clean the variable names, returning a data.frame:
```{r}
test_df %>%
  clean_names()
```
Compare to what base R produces:
```{r}
make.names(names(test_df))
```
