Inference Procedures
========================================================


### Exercise 1 - Bartlett Test for Homogenity of Variance

**Two Sample Case**
- this is equivalent to a test of equal variance for two samples.
- the null hypothesis is that both groups have the same variance for a particular numeric variable. 
- A boxplot is added to assist in interpretation.
- In this example, we fail reject the null hypothesis. The p-value is greater than 0.05.
```{r fig.width=7, fig.height=6}
boxplot(mpg~am,data=mtcars,horizontal=TRUE,col=c("lightblue","lightgreen"))
bartlett.test(mpg ~ am,data=mtcars)
```

### Exercise 2 - Bartlett Test for Homogenity of Variance

**Three Sample Case**
- The null hypothesis is that there is no significant difference in variance for a particular numeric variable between any of the groups.
- A boxplot is added to assist in interpretation.
- In this example, we reject the null hypothesis on account of the low p-value.
```{r fig.width=7, fig.height=6}
boxplot(mpg~cyl,data=mtcars,horizontal=TRUE,col=c("lightblue","lightgreen","lightpink"))
bartlett.test(mpg ~ cyl,data=mtcars)
```


