## glue_sql Interpolate strings with SQL escaping

#### Description
SQL databases often have custom quotation syntax for identifiers and strings which make writing SQL queries error prone and cumbersome to do. ``glue_sql()`` and ``glue_data_sql()`` are analogs to ``glue()`` and ``glue_data()`` which handle the SQL quoting.

#### Usage
<pre><code>
glue_sql(..., .con, .envir = parent.frame(), .na = DBI::SQL("NULL"))

glue_data_sql(.x, ..., .con, .envir = parent.frame(), .na = DBI::SQL("NULL"))
</code></pre>

Arguments
* ``...`` [expressions]:  Expressions string(s) to format, multiple inputs are concatenated together before
formatting.
* ``.con`` [DBIConnection]: A DBI connection object obtained from DBI::dbConnect().
* ``.envir`` [environment: parent.frame()] Environment to evaluate each expression in. Expressions are evaluated from left
to right. If .x is an environment, the expressions are evaluated in that environment
and .envir is ignored.
* ``.na`` [character(1): ‘NA’] Value to replace NA values with. If NULL missing values are propagated, that is
an NA result will cause NA output. Otherwise the value is replaced by the value
of .na.
* ``.x`` [listish] An environment, list or data frame used to lookup values.

#### Details
* They automatically quote character results, quote identifiers if the glue expression is surrounded by backticks ‘ and do not quote non-characters such as numbers. If numeric data is stored in a character column (which should be quoted) pass the data to’glue_sql()‘ as a character.
* Returning the result with DBI::SQL() will suppress quoting if desired for a given value.
* Note parameterized queries are generally the safest and most efficient way to pass user defined
values in a query, however not every database driver supports them.
* If you place a * at the end of a glue expression the values will be collapsed with commas. This is
useful for the SQL IN Operator for instance.

#### Value
A ``DBI::SQL()`` object with the given query.

#### Examples
```{r}
con <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")
colnames(iris) <- gsub("[.]", "_", tolower(colnames(iris)))
DBI::dbWriteTable(con, "iris", iris)
var <- "sepal_width"
tbl <- "iris"
num <- 2
val <- "setosa"
glue_sql("
SELECT {`var`}
FROM {`tbl`}
WHERE {`tbl`}.sepal_length > {num}
AND {`tbl`}.species = {val}
", .con = con)
```

```{r}
# If sepal_length is store on the database as a character explicitly convert
# the data to character to quote appropriately.
glue_sql("
SELECT {`var`}
FROM {`tbl`}
WHERE {`tbl`}.sepal_length > {as.character(num)}
AND {`tbl`}.species = {val}
", .con = con)
```

```{r}
# `glue_sql()` can be used in conjuction with parameterized queries using
# `DBI::dbBind()` to provide protection for SQL Injection attacks
sql <- glue_sql("
SELECT {`var`}
FROM {`tbl`}
WHERE {`tbl`}.sepal_length > ?
", .con = con)
query <- DBI::dbSendQuery(con, sql)
DBI::dbBind(query, list(num))
DBI::dbFetch(query, n = 4)
DBI::dbClearResult(query)
```

```{r}
# `glue_sql()` can be used to build up more complex queries with
# interchangeable sub queries. It returns `DBI::SQL()` objects which are
# properly protected from quoting.

sub_query <- glue_sql(" SELECT *  FROM {`tbl`} ", .con = con)

glue_sql(" SELECT s.{`var`}FROM ({sub_query}) AS s ", .con = con)
```

```{r}
# If you want to input multiple values for use in SQL IN statements put `*`
# at the end of the value and the values will be collapsed and quoted appropriately.
glue_sql("SELECT * FROM {`tbl`} WHERE sepal_length IN ({vals*})",
vals = 1, .con = con)
```

```{r}
glue_sql("SELECT * FROM {`tbl`} WHERE sepal_length IN ({vals*})",
vals = 1:5, .con = con)

```

```{r}
glue_sql("SELECT * FROM {`tbl`} WHERE species IN ({vals*})",
vals = "setosa", .con = con)
glue_sql("SELECT * FROM {`tbl`} WHERE species IN ({vals*})",
vals = c("setosa", "versicolor"), .con = con)
DBI::dbDisconnect(con)
```
