
###  Dates and data.frames
If you want to store your date in a data.frame you will NOT be able to use POSIXlt. 
The reason for this is that a POSIXlt actually is a list with 9 elements. So if you want to add your dates to a data.frame you will need to convert your dates to POSIXct:

data$time = as.POSIXct(strptime(data$time_string, "%H:%M:%S"))

####Format a Date
format(d1, "%a %Y/%m/%d") #[1] "So 2008/05/18" format(d2, "%A %Y/%m/%d") # [1] "Dienstag 2008/01/01"

#### Time Intervals

```{r}
b1 <- ISOdate(1977,7,13) 
b2 <- ISOdate(2003,8,14) 
b2 - b1 # Time difference of 9528 days 
> class(b2-b1) 
[1] "difftime"
```
If an alternative unit of time is desired, the <tt>difftime</tt> function can be called, using the optional <tt>units=</tt> argument with any of the following values: "auto", "secs", "mins", "hours", "days", or "weeks".
difftime(b2,b1,units="weeks") #Time difference of 1361.143 weeks

#### Time Sequences
The by=> argument to the seq function can be specified either as a difftime value, or in any units of time that the difftime function accepts, making it very easy to generate sequences of dates.
```{r}
seq(as.Date("1976-07-04"),by="days",length=10) 
# 
[1] "1976-07-04" "1976-07-05" "1976-07-06" "1976-07-07" "1976-07-08" 
[6] "1976-07-09" "1976-07-10" "1976-07-11" "1976-07-12" "1976-07-13" 
seq(as.Date("2000-06-01"),to=as.Date("2000-08-01"),by="2 weeks") 
# 
[1] "2000-06-01" "2000-06-15" "2000-06-29" "2000-07-13" "2000-07-27" 
seq(as.POSIXct("2009-03-23 00:00:00", tz="GMT"), length=96, by="15 mins") 
# 
[1] "2009-03-23 00:00:00 GMT" "2009-03-23 00:15:00 GMT" 
[3] "2009-03-23 00:30:00 GMT" "2009-03-23 00:45:00 GMT"
```
#### Formulas
Formulas in R can be thought of as a "little language" since they obey a different structure and syntax from expressions. 
Expressions when evaluated produce some result such as a number, vector or list which is then displayed by the print function. 
Formulas on the other hand are used as a concise and intuitive way of specifying a statistical model. 
For example, consider a multiple linear regression of y on a numeric variable x1 and its squared value, x1^2 and a categorical variable x2. 
Note that in R categorical variables are called factors. This regression is specified by:
y ~ x1 + I(x1^2) + x2
and could be fit using the lm function (linear model, regression):
lm(y ~ x1 + I(x1^2) + x2)

In the formula notation, "~" means the left-hand-side is the independent variable or response and the right-hand-side are the dependent variables. 
The I(x1^2) means interpret the inside expression as a regular expression in R. Including a factor variable like x2 is very convenient since we don't have to be bothered about specifying all the indicator variables as we would have to do in other statistical software.

[http://cran.r-project.org/doc/manuals/R-intro.html#Formulae-for-statistical-models Defining statistical models; formulae]
