---
title: "Introduction to dplyr"
author: "Daniel McInerney, Kevin O'Brien"
date: "18 February 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction to dplyr

This R markdown file is an introduction to data analysis and visualisation using tidyverse, specifically dplyr and ggplot2:

The exercise uses a subset of the Sub-compartment database, which has been exported as a comma separated file (CSV). 

### Getting Started
Let's start by reading in the data using `read.csv`. Note that you can specify if the file has a header and what the field separator is.
```{r}
scdb <- read.csv('../DATA/scdb_subset.csv', head=TRUE, sep=',')
```


Check that the data has been correctly assigned to the scdb object

```{r}
ls()
```

Review the internal structure of the scdb object.  
```{r}
str(scdb)
```

Note that first line of the output, what data type is it?

Do you know a function that can be used to confirm this data type?


You can show the full contents of the object by simply typing `scdb`:
```{r, eval=FALSE}
scdb
```

Return the field names associated with the object
`scdb`:
```{r}
names(scdb)
```

### Summarising your data

And summarise the data stored in the object
```{r, eval=TRUE}
summary(scdb)
```

Extract subsets for summaries. Firstly for a specific field, note that you identify a field using the "$" and then the field name. Secondly, select the first five columns, using [1:5]

```{r}
summary(scdb$SURVEY_YEA)

summary(scdb[1:5])
```

Note, summarise only the first 30 rows, note the use of the comma:

```{r}
summary(scdb[1:30,])
```



## Getting Started with dplyr

load dplyr package
```{r}
library(dplyr)
```

We will present some of the key dplyr functions that facilitate data manipulation in R. These include:

1. **select()** - pick variables by their names
2. **filter()** - pick observations by their values
3. **arrange()** - reorder the rows
4. **mutate()** - create new variables with functions using existing variables
5. **summarize()** - collapse many values down to a single summary

One unique feature of dplyr is the 'pipe' function "%>%", which can be used to chain functions together.

### select() function


Let's start by selecting 3 columns (area, plant year and species) to a new object (scdb_subset):
```{r}

scdb %>% 
    select(gis_area, pyr1, spp1) -> scdb_subset
summary(scdb_subset)

```

You could do this in one step as follows:
```{r}

scdb %>% 
  select(gis_area, pyr1, spp1) %>% 
  summary()

```



More flexible select statements, for instance select all the species fields (i.e. SPP1, SPP2, SPP3, ...) using the **contains()** helper function:

```{r}

scdb %>% 
  select(com_sub, gis_area, contains('spp')) -> scdb_subset_species

names(scdb_subset_species)
```


Let's suppose, you want to omit Species 4 from the above command:

```{r}

scdb %>% 
  select(com_sub, gis_area, contains('spp'), -(spp4)) %>%
  summary()

```

Yet another way to achieve the same thing using the **num_range()** helper function

```{r}

scdb %>% 
  select(com_sub, gis_area, num_range('spp',1:3)) %>%
  summary()

```

**Exercise**: What happens if you include a variable name twice in the select function?


Let's suppose that you want to move the above variables to the start of the data frame, than you can use the **everything()** function as follows:

```{r, eval=FALSE}

scdb %>% 
  select(com_sub, gis_area, num_range('spp',1:3), everything()) %>%
  summary()

```



**Exercise**: Select all fields, but exclude all species fields



### filter() function

**filter()** - pick observations by their values using specific expressions. The filter function is synonymous with queries in SQL. 



Let's filter our data frame to retrieve all rows where the first species is Sitka spruce (SS):

```{r}

scdb %>% 
  filter(spp1 == 'SS') %>% 
  summary() 

```


Let's filter our data frame to retrieve all subs were the first species is Sitka spruce (SS) and where they are older than 30 years of age:

```{r}

scdb %>% 
  filter(spp1 == 'SS' & age1 > 30) %>% 
  summary() 

```

**Exercise**: From what you learned this morning, what function could you use to determine how many records are in your result?




### arrange() function

Order columns using the *arrange()* function. 

Let's select some fields from our data frame and order them by area. 

### 'arrange' function - to sort your data 
```{r}

scdb %>% 
  select(compartmen, subid, gis_area) %>%
  arrange(gis_area) %>% 
  head()

```

### Combining select() and filter()

**Exercise**: Using the SCDB, select all subs where the first species is Lodgepole and that were planted before 2010 and summarise the area 





