---
title: "The Survival Functions"
subtitle: "Survival Analysis with R"
author: DragonflyStats.github.io
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=3)
```

### Example Datasets

This workshop will use a number of different time-to-event datasets, so we
look at those first.

#### Telco Churn Data

The telco churn data contains some customer usage information on mobile phone
customers and whether or not they left their subscription accounts.

```{r show_telco_churn_data, echo=TRUE}
telco_data %>% glimpse()

telco_data %>% head()
```


#### Prostate Survival

The prostate survival data is generated data from a cancer survival study.

```{r show_prostate_data, echo=TRUE}
prostate_surv_tbl %>% glimpse()

prostate_surv_tbl %>% head()
```

### Data Setup Script
library(tidyverse)
library(asaur)

# Re-using some code from Sec 1 of the workshop to avoid repeating myself
# in later worksheets.


# Setup the Telco churn data
telco_cols <- cols(
    areacode = col_character()
)

telco_churn_tbl <- read_csv('data/telcochurn.csv'
                           ,col_types = telco_cols
                            )

# Setup the pharmacoSmoking data
pharmaco_smoker_tbl <- asaur::pharmacoSmoking %>% as_data_frame()

# Setup the prostate survival data
prostate_surv_tbl <- asaur::prostateSurvival %>% as_data_frame()