---
title: "``cbreaks`` Compute breaks for continuous scale."
author: "Kevin O'Brien"
date: "23 November 2018"
output: html_document
---

```{r}
library(scales)
```
#### Description

This function wraps up the components needed to go from a continuous range to a set of breaks and
labels suitable for display on axes or legends.

#### Usage
<pre><code>
cbreaks(range, breaks = extended_breaks(),
  labels = scientific_format())
</code></pre>

#### Arguments
* ``range``: numeric vector of length 2 giving the range of the underlying data
* ``breaks``: either a vector of break values, or a break function that will make a vector of breaks when given the range of the data
* ``labels``: either a vector of labels (character vector or list of expression) or a format function
that will make a vector of labels when called with a vector of breaks. Labels can only be specified manually if breaks are - it is extremely dangerous to supply labels if you don’t know what the breaks will be.

#### Examples

```{r}
cbreaks(c(0, 100))
cbreaks(c(0, 100), pretty_breaks(3))
cbreaks(c(0, 100), pretty_breaks(10))
```


```{r}
cbreaks(c(1, 100), log_breaks())
cbreaks(c(1, 1e4), log_breaks())
```


```{r}
cbreaks(c(0, 100), labels = math_format())
cbreaks(c(0, 1), labels = percent_format())
cbreaks(c(0, 1e6), labels = comma_format())
```


```{r}
cbreaks(c(0, 1e6), labels = dollar_format())
cbreaks(c(0, 30), labels = dollar_format())
```

```{r}
# You can also specify them manually:
cbreaks(c(0, 100), breaks = c(15, 20, 80))
```


```{r}
cbreaks(c(0, 100), breaks = c(15, 20, 80), labels = c(1.5, 2.0, 8.0))
cbreaks(c(0, 100), breaks = c(15, 20, 80),
labels = expression(alpha, beta, gamma))
```
