---
title: "``cscale``: Continuous scale."
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(scales)
```
 

#### Description
Continuous scale.

#### Usage
<pre><code>
cscale(x, palette, na.value = NA_real_, trans = identity_trans())
</code></pre>

#### Arguments
* ``x`` vector of continuous values to scale
* ``palette``: palette to use.
* Built in palettes: area_pal, brewer_pal, dichromat_pal, div_gradient_pal, gradient_n_pal, grey_pal, hue_pal, identity_pal, linetype_pal, manual_pal,
rescale_pal, seq_gradient_pal, shape_pal, viridis_pal
* ``na.value`` value to use for missing values
* ``trans``: transformation object describing the how to transform the raw data prior to scaling.


#### Examples

```{r}
with(mtcars, plot(disp, mpg, cex = cscale(hp, rescale_pal())))
```

```{r}
with(mtcars, plot(disp, mpg, cex = cscale(hp, rescale_pal(),
                                          trans = sqrt_trans())))
```

```{r}
with(mtcars, plot(disp, mpg, cex = cscale(hp, area_pal())))
```

```{r}
with(mtcars, plot(disp, mpg, pch = 20, cex = 5,
                  col = cscale(hp, seq_gradient_pal("grey80", "black"))))

```

Defaults to the identity transformation which leaves the data unchanged.
Built in transformations: asn_trans, atanh_trans, boxcox_trans, date_trans,
exp_trans, hms_trans, identity_trans, log10_trans, log1p_trans, log2_trans,
log_trans, logit_trans, modulus_trans, probability_trans, probit_trans,
pseudo_log_trans, reciprocal_trans, reverse_trans, sqrt_trans, time_trans.