Grouping, loops and conditional execution
R is an expression language in the sense that its only command type is a function or expression which returns a result. Even an assignment is an expression whose result is the value assigned, and it may be used wherever any expression may be used; in particular multiple assignments are possible.

Commands may be grouped together in braces, $\{expr_1; ...; expr_m\}$, in which case the value of the group is the result of the last expression in the group evaluated. Since such a group is also an expression it may, for example, be itself included in parentheses and used a part of an even larger expression, and so on.  


\subsection*{if Statement}

Have already encountered implicit looping when using the apply
family of functions.

Conditional execution: the if statement has the form 
\begin{verbatim}
if (condition){ # Brackets can be omitted if only one command
expr_1 # to be carried out.
}
else {
expr_2
}    
\end{verbatim}


The condition must evaluate to a logical value, i.e. TRUE or
FALSE. If the condition == TRUE, expr\_1 is carried out, which
can consist of a single command or multiple commands. If the
condition == FALSE, expr\_2 is carried out.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
Can also have longer if statements:
\begin{verbatim}
if (condition1){
expr_1
}
else if (condition2){
expr_2
}
...
else {
expr_n
}   
\end{verbatim}

If condition1 == TRUE, expr\_1 is executed and the checking
stops. If condition1 == FALSE, moves on to condition2 and
checks if that condition is met. If condition2 == TRUE, expr\_2
is executed and checking stops. If condition2 == FALSE, moves
on to the next condition and so on until all conditions have been
checked.
The final else is executed if none of the previous conditions have
returned a value of TRUE.


Usually the logical operators &&, ||, ==, !=, >, <, >=, <= are used
as the conditions in the if statement.
The following function gives a demonstration of the use of
if... else.
\begin{verbatim}
comparisons1 <- function(number)
{
# if ... else
if (number != 1)
{
cat(number,"is not one\n")
}
else
{
cat(number,"is one\n")
}
}
> comparisons1(1) > comparisons1(20)
1 is one 20 is not one
\end{verbatim}


The following demonstrates the use of
if ... else if ... else
\begin{verbatim}
 comparisons2 <- function(number)
{
if (number == 0)
{
cat(number,"equals 0\n")
}
else if (number > 0)
{
cat(number,"is positive\n")
}
else
{
cat(number,"is negative\n")
}
}
> comparisons2(0) > comparisons2(-15) > comparisons2(1)
0 equals 0 -15 is negative 1 is positive   
\end{verbatim}



This function demonstrates the use of && in the condition. This
means that both conditions must be met before a value of TRUE is
returned.
\begin{verbatim}
comparisons3 <- function(number)
{
if ( (number > 0) && (number < 10) )
{
cat(number,"is between 0 and 10\n")
}
}
> comparisons3(-1) > comparisons3(9) > comparisons3(10)
9 is between 0 and 10    
\end{verbatim}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\subsection*{ifelse Statement}
A vectorised version of the if statement is ifelse. This is useful
if you want to perform some action on every element of a vector
that satis es some condition.
The syntax is
ifelse( condition, true expr, false expr )
If condition == TRUE, the true expr is carried out. If
condition == FALSE, the false expr is carried out.
\begin{framed}
\begin{verbatim}
x <- rnorm(20, mean=15, sd=5)
x
[1] 23.608513 14.424667 12.306040 14.291568 18.522846 14.514071 22.004400
[8] 24.658249 11.697999 16.344976 22.110389 8.455789 19.672274 22.393680
[15] 11.449034 17.288859 14.839597 14.484774 18.636589 22.670548
ifelse(x >= 17, sqrt(x), NA)
[1] 4.858859 NA NA NA 4.303818 NA 4.690885
[8] 4.965707 NA NA 4.702169 NA 4.435344 4.732196
[15] NA 4.157987 NA NA 4.317012 4.761360
    
\end{verbatim}
\end{framed}


\subsection*{for Loops}
Repetitive execution: for loops, while loops and repeat loops.
To loop/iterate through a certain number of repetitions a for loop
is used. The basic syntax is
\begin{verbatim}
for(variable_name in sequence) {
command
command
command
}
\end{verbatim}

A simple example of a for loop is:
\begin{verbatim}
for(i in 1:5){
print(sqrt(i))
}
[1] 1
[1] 1.414214
[1] 1.732051
[1] 2
[1] 2.236068
\end{verbatim}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
Another example is:
\begin{verbatim}
n <- 20
p <- 5
value <- vector(mode="numeric", length=n)
rand.nums <- matrix(rnorm(n*p), nrow=n)
for(i in 1:length(value)){
value[i] <- max(rand.nums[i,])
print(sum(value))
}
\end{verbatim}
The first four lines create variables n and p with values 20 and 5
respectively, a numeric vector called value with length 20 and a
matrix of 20*5=100 random numbers, called rand.nums, with 20
rows.
The for loop performs 20 loops and stores the maximum value
from each row of rand.nums into position i of the vector value.
The sum of the current numbers in value is also printed to the
screen.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Can also have nested for loops. Indenting your code can be useful
when trying to \match" brackets.
for(variable_name1 in sequence) {
command
command
for(variable_name2 in sequence) {
command
command
command
} # ends inner for loop
} # ends outer for loop
It should be noted that variable_name2 should be di erent from
variable_name1, e.g. use i and j. Using the same name will
reset the counter each time and result in an in nite loop!!


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Load the function simple.nesting from Loops.R and call the
function using
simple.nesting(num.fam=5, num.child=3).
The le nest.dat will be created in your current working
directory. Open this file and explore the contents.
for loops and multiply nested for loops are generally avoided
when possible in R as they can be quite slow. We will use in
simulation examples later in the course.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The while loop can be used if the number of iterations required is
not known beforehand. For example, if we want to continue
looping until a certain condition is met, a while loop is useful.
The following is the syntax for a while loop:
\begin{verbatim}
while (condition){
command
command
}
The loop continues while condition == TRUE.
niter <- 0
num <- sample(1:100, 1)
while(num != 20) {
num <- sample(1:100, 1)
niter <- niter + 1
}
niter
    
\end{verbatim}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{next, break, repeat Statements}
The next statement can be used to discontinue one particular
iteration of any loop, i.e. this iteration is ended and the loop
\skips" to the next iteration. Useful if you want a loop to continue
even if an error is found (error checking).
The break statement completely terminates a loop. Useful if you
want a loop to end if an error is found. See the Loops.R script le
for code to exhibit the di erence between the next and break
statements.
The repeat loop uses next and break. The only way to end this
type of loop is to use the break statement. For an example, see
the Loops.R script file.
\end{document}
