
CS1B Paper I {data-navmenu="R for Actuarial Statistics"}
==============================================================

Column {.tabset}
--------------------------------


### Question 1
<h4> Question 1</h4>

```{r include=FALSE, warning=FALSE}
library(tidyverse)
library(data.table)
library(knitr)
library(magrittr)
library(flextable)
library(kableExtra)
```

In a small country, a political election was held recently to decide on a political party 
to govern the country. A survey was conducted to monitor the approval rating for the 
winning political party. 

A sample of 15 voters were asked to complete the same survey twice: once before the 
election (Approval before) and a second time 6 months after the election (Approval 
after). 

---

The survey asked each of the 15 voters to record their approval for the winning political party on a scale from 1 to 10, where 
1 = Strongly disapprove and 10 = Strongly approve. 
The results are shown in the table below:


```{r echo=FALSE, warning=FALSE}
voters <- as.character(1:15)

approval_before <- c(8, 6, 8, 7, 7, 4, 2, 10, 8, 7, 10, 
8, 8, 9, 6) 
approval_after <- c(5, 6, 2, 3, 4, 1, 4, 7, 4, 2, 10, 1, 
5, 3, 7) 

thisTable <- data.frame(voters,approval_before,approval_after)

thisTable_t <-thisTable %>% data.table::transpose() 

thisTable_t <- bind_cols(varnames=c("Voter","Before","After"),thisTable_t)

thisTable_t %>% kable( col.names = NULL) %>%  
      kable_styling(bootstrap_options = c("bordered", full_width = F, font_size = 16)) %>%
  row_spec(1, background = "AAEEFF", color = "black", font_size = 18)


```


The values can be entered in R using: 

<pre><code>
approval_before <- c(8, 6, 8, 7, 7, 4, 2, 10, 8, 7, 10, 8, 8, 9, 6)
 
approval_after <- c(5, 6, 2, 3, 4, 1, 4, 7, 4, 2, 10, 1,  5, 3, 7) 
</code></pre>

---


#### Exercises

1. Calculate the means for the Approval before and Approval after results. 

2.  Outline why Kendall’s Tau would be a suitable correlation coefficient to use 
for these results. 

3. Calculate Kendall’s Tau coefficient between the Approval before and 
Approval after results. 

4. Comment on your results in Exercises 1 and 3. 


### Question 2
<h4> Question 2</h4>
Use the command <tt>set.seed(1234)</tt> to initialise the random number generator. 


#### Exercises
1.  Generate a sample of size $n = 180$ from an Exponential distribution 
with parameter $\lambda  = 5$. 

2. Calculate the sample median of the sample in part 1.

3. Calculate the median of the Exponential distribution with $\lambda  = 5$. 

4. Comment on your results in parts 2 and 3. 

5.  The maximum likelihood estimator for $\lambda$ is given by $\lambda = 1/X$ where $X$ is the sample 
mean. Estimate the parameter $\lambda$  using the maximum likelihood estimator and 
the sample in part 1

6. Generate another sample of size $n = 180$ from an Exponential 
distribution with parameter $\lambda  = 5$ and estimate the parameter $\lambda$  using 
the maximum likelihood estimator and the new sample. 

7. Comment on the estimated values in parts 5 and 6



### Question 3
<h4> Question 3</h4>
An insurer’s marketing team has developed a new lottery that gives each new 
customer the chance to win a cash prize. The insurer states that any new customer can 
win with probability 0.36 independently of all other customers.

Let $Y$ denote the number of winners in a random sample of 900 new customers. 

#### Exercises

1.  State a suitable distribution for the random variable Y. 
2. Calculate, to four decimal places, the probabilities $P[k < Y \leq k + 20]$ for all 
values $k = 220, 240, 260,\ldots, 400$.

3. It is suggested that the true distribution of $Y$ can be approximated with a Normal 
distribution with expectation EሾYሿ = 324 and a standard deviation of 14.4. 
Calculate, to four decimal places, the probabilities $P[k < Y \leq k + 20]$ for all 
values $k = 220, 240, 260,\ldots, 400$ based on the suggested Normal distribution. 

4. Plot the values calculated in parts 2. and 3. on the same graph. 
[Hint: you may find the R command ***lines(…)*** useful.] 

5.Discuss your answers in parts 1 to 4.. 

### Question 4
<h4> Question 4</h4>

An actuarial modeller in a health insurance company is constructing a Generalised 
Linear Model (GLM) to analyse claim numbers for its critical illness policies. For 
every policy over the past year, the company has collected the number of reported 
claims (***Claim_number***) and data on the following covariates:

* ***Age***: Age of policyholder, a  number between 18 and 80 
* ***Gender***: Gender of the policyholder (Male or Female) 
* ***Region***: A description of the region where the policyholder lives 
* ***Pre_existing_health_condition***: A categorical valuerepresenting whether the 
policyholder has a pre-existing health condition, 0 = no and 1 = yes
<p>
The data given in the file named ***Claims_Experience.RData*** show the past year’s reported claims for this company’s critical illness policies. After loading the 
data into R, using the command load("Claims_Experience.RData"), the 
data frame ***data_claims*** with its variables listed above will be available. 

The modeller wants to fit a GLM with the Claim_number as the response variable 
and is deciding whether to fit the GLM using a Poisson or Normal distribution. 

#### Exercises

1.  Fit a GLM, using a Normal distribution, that treats age as a numerical variable and the remaining three covariates as factors, calling this model_g. Your 
answer should include the estimated value, standard error and p-value of each parameter in the model. 
2. 


<ol type="a">
<li> Fit a GLM, using the Poisson distribution, that treats age as a numerical variable and the remaining three covariates as factors, 
calling this ***model_p***. Your answer should include the estimated value, standard error and 
p-value of each parameter in the model.  </li>
<li> Justify which model the modeller should use, using your answers to 
parts 1.  and 2.(a) </li>
<li> standard deviation of premium </li>
<li> Comment on the dependence of the number of reported claims on 
Pre_existing_health_condition based on your answers to 
parts 2.(a) and 2.(b).  </li>
</ol>

The modeller has fitted the GLM using a Poisson distribution and now wants to 
investigate which of the three factors should be selected in the model, by using a 
method similar to the backward selection approach. The following models are 
considered: 
* model_p2: removes Gender from model_p
* model_p3: removes Gender and Pre_exisiting_health_condition from model_p
* model_p4: removes Gender, Pre_exisiting_health_condition and  Region from model_p 

3.  Compare the values of the Akaike’s Information Criterion for the four models 
model_p, model_p2, model_p3 and model_p4. 
4.  (a) Comment on your answer to part 3. . 
(b) Comment on the differences between the process in part 3.  and the full backward selection method. 

The modeller suggests refining the fitted GLM (model_p) by also including the 
interactions between the variables: Age, Region and 
Pre_existing_health_condition. 

5.  (a) Fit an appropriate model, to include the interactions between the 
variables Age, Region and 
***Pre_existing_health_condition***. 
(b) Justify whether the refined model in part 5. (a) improves the model 
(model_p) fitted in part 2.(a). 
(c) Comment on whether any of the interactions are significantly  associated with claim numbers. 

### Question 5
<h4> Question 5</h4>

An analyst collected samples of the prices (in £000s) of 25 one-bedroom flats in each of two different cities. The data from these two samples are saved in the file 
onebedflat.Rdata. After loading the data into R, using the command ***load("onebedflat.Rdata")***, the data frame onebedflat with its variables 
listed below will be available. 

* City1: one-bedroom flat prices in the first city. 
* City2: one-bedroom flat prices in the second city. 

It is assumed that the two samples come from Normal populations with equal 
variances. 

#### Exercises

1. Calculate an appropriate test statistic for the hypothesis of equal means in the two corresponding populations using these data.  
2. Test at a 1% significance level whether the mean flat prices are the same in the two cities, against the alternative that they are different, based on your answer 
to part 1. . 
The standard assumptions for a two-sided test of the hypothesis of equal means in two 
populations are that the populations: 

* (A1) follow Normal distributions 
* (A2) have equal variances. 
The analyst suspects that those assumptions may not be satisfied, and they are therefore interested in learning about the true significance level for such tests when 
assumptions (A1) and (A2) do not necessarily hold. 
With two independent samples, given a particular choice of the significance level, \alpha, 
the true significance level for a two-sided test of the hypothesis of equal means can be calculated as 
\alphatrue ൌ 𝑃ሺ|test statistic| ൒ critical pointሻ,
where the critical point is the \alpha/2 quantile of an appropriate distribution. 


An estimate of the true significance level for this two-sided test, when the chosen 
level of significance is $\alpha = 0.1$, can be computed using the R Monte Carlo simulation 
code provided below: 
<pre><code>
set.seed(123) # 
alpha = 0.1 
m=20 
n=20 
N=8000 
nrej=0 
for (i in 1:N) { # 
x=rnorm(m,mean=0,sd=1); y=rnorm(n,mean=0,sd=1) # 
ts=t.test(x, y, var.equal=TRUE)$stat # 
df = n+m-2 # 
if (abs(ts)>=qt(1-alpha/2, df)) # 
nrej=nrej+1 # 
} 
alpha_est=nrej/N # 
</code></pre>
Note that the code provided above lacks best practice of including appropriate 
comments. 
3.  Explain the provided code by placing comments after each ‘#’ sign in 
the code. 

4.  Estimate the true significance level, by using the code provided above, when the two populations are from a standard Normal distribution. 
5.  Estimate the true significance level, by adjusting the code provided above to reflect the following pairs of populations: 

<ol type="a">
<li> The first population is from a standard Normal distribution and the 
second population is from a Normal distribution with mean 0 and standard deviation 10.</li>
<li> The first population is from a Normal distribution with mean 10 and standard deviation 2 and the second population is from an Exponential 
distribution with mean 10.  </li>
</ol>


6.  Compare your estimates of the true significance level under all scenarios in 
parts 4. , 5. (a) and 5. (b), stating your conclusions. 
