
CS2B Paper I {data-navmenu="R for Risk Modelling and Survival Analysis"}
==============================================================

Column  {.tabset}
--------------------------------

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

### Question 1
<h4> Question 1</h4>
The ‘***CS2B_Sept_21_Qu_1_Data.csv***’ file contains mortality data for a particular population of females in the year 1921, taken from the Human Mortality Database. 
The file contains the following three variables:

* ***Age*** Age, 𝑥, in single years from 25 to 90 
* ***mu_x*** Crude values of the forces of mortality at age x, μ௫
* ***Exposed_x*** Exposed to risk at age $X$ nearest birthday 

Before answering this question, the ‘CS2B_Sept_21_Qu_1_Data.csv’ file should be 
loaded into R and assigned to a data frame called mortalitydata. 

#### Exercises

1. Plot the crude values of μ௫ against age, x, using a line graph. [

2. Comment on the age pattern of the crude values of μx
 shown in the graph in exercise 1part (i). 

--------------

The Gompertz law of mortality can be expressed as follows: 
μx
 = Bcx

3. Perform a graduation of the crude values of μ௫ using the Gompertz formula, 
specifying the values of B and c in your answer script. 
4. Perform a chi-square goodness-of-fit test of the graduated values of μx
 from exercise 3 using a 95% confidence level. 
 
### Question 2
<h4> Question 2</h4>

2 Before answering this question, the R package for copulas should be loaded into R 
using the following code: 
```{r eval=FALSE}
install.packages("copula") 
library(copula) 
```

A copula model is to be constructed of the dependence between annual average 
temperature X, in degrees Celsius, and annual rainfall Y, in millimetres, in a region at 
risk of desertification. The marginal distributions of X and Y are: 
X ~ N൫20, 32
൯, Y ~ N൫200, 502
൯. 
----

Two copulas are under consideration to model the dependence of X and Y: 
* Copula 1: A Gaussian copula with ρ = –0.5. 
* Copula 2: A Student’s t copula with 3 degrees of freedom and ρ = –0.5. 

The following R code generates n simulations of jointly distributed random variables U and V, where the marginal distributions of U and V are the uniform distribution on 
[0, 1] and the dependency between U and V is given by a Gaussian copula with correlation ρ (rho): 
```{r eval=FALSE}
rCopula(n, normalCopula(param = rho)) 
```
The corresponding code for the Student’s t-copula with nu degrees of freedom is: 
```{r eval=FALSE}
rCopula(n, tCopula(param = rho, df = nu)) 
```
The output from the function rCopula is in the format of an n × 2 matrix whose rows are the observations of (U, V). 

Before answering exercise 1, generate the matrix, GaussXY, in R using the following code: 
```{r eval=FALSE}
set.seed(3) 
GaussUV <- rCopula(200000, normalCopula(param = -0.5)) 
GaussXY <- matrix(nrow = 200000, ncol = 2) 
GaussXY[,1] <- qnorm(GaussUV[,1], mean = 20, sd = 3) 
GaussXY[,2] <- qnorm(GaussUV[,2], mean = 200, sd = 50) 
```

The 200,000 × 2 matrix, GaussXY, represents 200,000 observations of (X, Y) under 
Copula 1. 

#### Exercises
1. Generate 200,000 observations of (X, Y) using Copula 2 and the marginal distributions of X and Y, assigning the observations to a 200,000 × 2 matrix 
named t3XY. You should set a random number generator seed of 3 before generating the observations and use the R function, ``head()``, to display the 
first six rows of t3XY in your answer script. 
2. Plot, on a single graph, line graphs of the estimated conditional mean of Y
given that X ൒ x against x for values of x at intervals of 1 from 10 to 30 
inclusive for each of the two generated copulas. You should use separate 
colours to identify each of the two line graphs and set the y-axis range from 50 
to 250. 
3. Comment on the key features of your plot in exercise 2.  
4. Comment on the implications of your plot in exercise 2 if the distribution of 
temperatures, X, is expected to become more weighted towards higher values 
as a result of climate change.  

### Question 3
<h4> Question 3</h4>

Before answering this question, the R packages for calculating and plotting Recursive 
Partitioning and Regression Trees should be loaded into R using the following code: 
```{r eval=FALSE}
install.packages("rpart") 
install.packages("rpart.plot") 
library(rpart) 
library(rpart.plot) 
```

A new start-up bank is considering which product it should first launch to market. 
One of the products it is considering is a personal loan. An Actuary employed by the 
bank wants to construct a simple decision tree machine learning model to determine 
which loan applications to approve. 

The Actuary has decided to construct the decision tree model to classify each 
potential customer as either expected to default (default = 1) or not expected to default 
(default = 0) based on the following inputs: 
*  a categorical feature, f1, that takes values –1, 0 or +1 
*  a continuous feature, f2, that takes values between –1 and 1. 
With no previous customer data to draw on, the Actuary decides to construct a 
specimen data set to train the decision tree model containing the following 
information for each of n hypothetical customers, where n = 10,000: 
*  feature f1 taking the values –1, 0 and +1 with equal probability 
*  feature f2 being uniformly distributed on [–1,1] 
*  a field containing the value 1 if the customer defaults on the loan and 0 otherwise. 
To determine whether a hypothetical customer will default, the Actuary decides to 
model the probability of default as: 
exp (f)
1 + exp (f)
where f = a + bf1 + cf2 and a, b and c are parameters. 
Before constructing the specimen data set, the Actuary first needs to generate three 
sets of 10,000 observations from the uniform distribution on [0, 1]. 

#### Exercises

1. Generate a 10,000 × 3 matrix named U of observations from the uniform distribution on [0, 1]. 
Each column of U should contain 10,000 observations that have been generated together but separately from the observations in other columns. 
You should set a random number generator seed of 12345 before generating the observations in the first column and use the R function, 
head(), to display the first six rows of U in your answer script. 

Let Uij represent the (i, j) entry of matrix U. 
The Actuary proceeds with creating the specimen data set by constructing: 
*  an n-component vector f1 whose ith component is equal to –1 if Ui1 ≤ 1
3
, to 
0 if 1
3 < Ui1 ≤ 2
3
 and to +1 if Ui1 > 2
3
*  an n-component vector f2 whose ith component is equal to 2(Ui2 – 0.5) 
*  an n-component vector f whose ith component is the value of f given by the formula above for customer i
*  an n-component vector defprob whose ith component is the probability of default given by the formula above for customer i
*  an n-component vector default whose ith component is equal to 1 if Ui3 is less than or equal to the probability from the vector defprob for customer i and 0 otherwise 
*  a data frame in the format required to construct decision trees using the R package ***rpart***, incorporating the vectors f1, f2 and default. 
<p>
The R code to create the data frame in the format required, given the vectors f1, f2 and 
default, is: 
data.frame("f1" = f1, "f2" = f2, "default" = default)
<p>
2. Generate a data frame named specimen, in line with the Actuary’s construction 
rules, corresponding to the parameter values n = 10,000, a = 0 and b = c = 0.5. 
Use the R function, head(), to display the first six rows of specimen in your 
answer script. 

3. Determine the expected probability of customer default based on the specimen 
data. 

4. Comment on how realistic your answer to exercise 3 is in terms of what the bank may expect in practice, suggesting how the parameter a could be 
changed to make it more realistic. 
The R code for fitting and plotting a decision tree object called tree, that predicts whether a customer will default or not, is as follows: 
<pre><cpde>
tree = rpart(default ~ ., data = specimen, method = "class")
rpart.plot(tree, digits = 3, type = 5, extra = 106) 
```

5. Plot the decision tree object, called tree, using the R code above. 
6. State, using the decision tree plot, whether a customer with features f1 = 0 and 
f2 = –0.425 is expected to default or not. 

### Question 3 Part (b)
<h4> Question 3</h4>

Before launching the product, the Actuary gathers actual customer data from another 
bank, that is available in the public domain, to assess the predictive power of the decision tree model. 

The ***CS2B_Sept_21_Qu_3_Data.csv*** file contains the actual data for 100 customers in the required format for the decision tree model. 
Before answering the remainder of this question, the ‘CS2B_Sept_21_Qu_3_Data.csv’ file should be loaded into R and assigned to a data frame called actual. 

#### Exercises

7. Show the first six rows of the actual data frame in your answer script using the 
R function, head(). 
The R code for generating the predicted default classifications of the policyholders in the actual data frame using the tree decision tree object is as follows: 
```{r eval=FALSE}
predict(tree, actual, type = 'class') 
```

8. Generate the predicted default classifications of the 100 customers in the actual data frame using the R code above, assigning the predictions to a vector called predict_defaults. Use the R function, ``head()``, to display the first six values of predict_defaults in your answer script. 

9. Generate a confusion matrix of actual defaults versus predicted defaults for the 
100 actual customers, displaying the matrix in your answer script.  

10. Determine, using R, the values of the precision and recall percentages for the 
decision tree model’s predictions for the 100 actual customers, where the true positive is defined as the case where the model predicts a default for a 
customer that has actually defaulted. 

Based on the value of the recall percentage for the 100 actual customers, the bank’s 
Finance Director concludes that the decision tree model needs no further refinement 
and can be used to determine which loan applications to approve when the product is 
launched. 

11. Comment on the Finance Director’s conclusion. 
