
CS2B Paper A {data-navmenu="R for Risk Modelling and Survival Analysis"}
==============================================================

Column {.tabset}
--------------------------------


### Question 1
<h4> Question 1</h4>
Data is available for the movement of taxis in Dublin. 

The city is divided into three zones “North”, “South” and “West”. (*The “East” is Dublin Bay*)

The movement of a taxi from one zone to another will depend only on its current position. 

The following probabilities have been determined for taxi movements:


* Of all taxis in the North zone, 30% will remain in North and 30% will move to South, with the remaining 40% moving to West.

* In the South zone, taxis have a 40% chance of moving to North, 40% chance of staying in South and 20% chance of moving to West.

* Of all the drivers in the West zone, 50% will move to North and 30% to South with the remaining 20% staying in West.

#### Driver Movement
The movement of taxis in Dublin will be modelled in R using a Markov Chain.


A 3 - dimensional discrete Markov Chain defined by the following states:
North, South, West

The transition matrix (by rows) is defined as follows:

|| North | South | West |
| :----:| :----:|:----:|:----:|
|North | 0.3 | 0.3 | 0.4|
|South | 0.4 | 0.4 | 0.2|
|West | 0.5 | 0.3 | 0.2|

### Question 1 Exercises

### Question 2

<h4> Question 2  </h4>
* An exponential distribution has a parameter of $\lambda = 0.4$.

* Use the in-built functions in R to perform the following tasks.



#### Exercise 1


Simulate 1,000 values from this distribution, assigning this to a variable called ***Exp_Vector*** and calculate the mean and variance of
the simulated values. Paste the results of your calculation into your answer.



#### Exercise 2
Plot a histogram of Exp_Vector showing the frequencies and paste 
the plot into your answer.


#### Exercise 3
Plot the probability density function for this distribution as:

a. a scatter plot
b. a line graph.
Paste your plots into your answer
 
 
 
### Question 3 - Part (a)

<h4>Question 3 - Part (a)</h4>
You have been asked to analyse the percentage changes in quarterly personal 
consumption expenditure and personal disposable income in the US from 1970–2010. 
This information is contained in a time series called “consumption” in R’s fpp library.
The data can be loaded into R using the command: 

```{r eval=FALSE}
install.packages(“fpp”)
library(fpp)

consumption <- usconsumption[,1]
```


#### Exercise 1
Plot this time series giving appropriate labels for each axis and paste the R 
code and the chart into your answer. 

#### Exercise 2
Plot the Autocorrelation function (ACF) and Partial Autocorrelation 
function (PACF), giving appropriate labels for each axis and paste the R code and the charts into your answer. <p>
Comment on your plots making reference to the stationarity. 

#### Exercise 3
Fit the most appropriate ARIMA model based on the results in exercise 2, 
stating the equation of the model and justifying your choice. 


### Question 3 (Part (b))

The dataset usconsumption also includes the quarterly personal disposable income from 1970 to 2010. 
<p>
The data can be loaded into R using the command:

```{r eval=FALSE}
income <- usconsumption[,2]
```

#### Exercise 4 
Compare the performance of the ARIMA model you have chosen in 
exercise 3 with a linear regression model of consumption on income, by computing the root mean square error (RMSE) for the fitted values ofeach model. Paste your coding and R output into your answer.
<p>
Explain why the RMSE is not the ideal measure to compare these models, including a recommendation of a more appropriate measure.

<p>
#### Exercise 5
An analyst has suggested that neither model in exercise 4 is a good fit to the data and has asked you to propose an alternative model.

* Suggest a suitable alternative model to fit to the data. 
* Fit the model you have suggested to the data, stating the equation used.
* Compare the results of this model to the models fitted in exercise 4. 
 