CS2B Paper G {data-navmenu="R for Risk Modelling and Survival Analysis"}
==============================================================


```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


Column {.tabset}
--------------------------------

### Question 1

<h4>Question 1</h4>


1 Before answering this question, the R package for Markov chains should be loaded 
into R using the following code: 
```{r eval=FALSE}
install.packages("markovchain") 
library(markovchain)
```
$X_t$ is a Markov chain on the state space {1,2,3} with the following transition matrix: 
$$
\begin{array}{cccc}
p & 1 – p & 0\\
q & 0 & 1 – q\\
0 & r & 1 – r\\
\end{array}
$$
where $0 < p, q, r < 1$. 

#### Exercises

1.  Construct an R function, with arguments p, q and r, that generates a Markov chain object for $X_t$. 
 Assume that $p = 0.75$ and $r = 0.25$. 

2.  Construct R code that calculates the stationary distribution of $X_t$ for values of q from 0.1 to 0.9 inclusive, at intervals of 0.1, pasting your results into your 
answer script.<p> 

3. Plot a graph showing the stationary distributions of Xt calculated in exercise 2 as 
a function of q. You should include three separately coloured lines on your graph, each line representing the stationary probabilities of each state.  

### Question 2

<h4>Question 2</h4>


2 Before answering this question, construct the function, rpareto, in R using the 
following code: 

```{r eval = FALSE}
rpareto = function(n, alpha, lambda){ 
rp = lambda * ((1 - runif(n))^(-1/alpha) - 1) 
rp}
```
The rpareto function generates a random sample of size n from a two-parameter 
Pareto distribution with parameters \alpha and \lambda. 

#### Exercises

1. Generate, using rpareto, a random sample of size 25,000 from the 
two-parameter Pareto distribution with parameters $\alpha = 3$ and $\lambda = 1$, assigning 
the simulated values to a vector called <tt>A_vec</tt>. You should set a random number generator seed of 123 before generating <tt>A_vec</tt> and use the R function, 
<tt>head()</tt>, to display the first eight values of <tt>A_vec</tt> in your answer script.  

2. Construct an R function, called <tt>A_exceed_u</tt>, with two arguments, A and u, that 
returns only the non-zero entries of a vector, E, where vector E is of length 
25,000, with entries defined as: 
Ei = maxሼAi – u, 0ሽ.
Use the R function, <tt>head()</tt>, to display in your answer script the first eight 
values of <tt>A_exceed_u</tt> when A = <tt>A_vec</tt> and u = 1.  

3.  Construct another R function, called <tt>F_u</tt>, with one argument, A_greater_than_u, that returns a vector of length 101, containing the 
probabilities: 
PሺAi – u ≤ x | Ai > u), where x ≥ 0, 
for values of x from 0 to 10 inclusive, at intervals of 0.1. 
Use the R function, <tt>head()</tt>, to display in your answer script the first eight values of <tt>F_u</tt> when ***A_greater_than_u = <tt>A_exceed_u</tt>*** with arguments 
A = <tt>A_vec</tt> and u = 1. 

4. Plot, on a single graph, four line graphs of the values of <tt>F_u</tt> against x for the values of x specified in exercise 3 when A_greater_than_u = <tt>A_exceed_u</tt>, 
A = <tt>A_vec</tt> and u = 1, 2, 3 and 4. You should use separate colours to identify each line graph and you should clearly specify which value of u is represented 
by each of the four line graphs.  

5. Comment on the graph produced in exercise  (4. . 

### Question 3

<h4>Question 3</h4>
Before answering this question, the ‘survival’ package should be loaded into R using the following code: 
```{r eval=FALSE}
install.packages("survival") 
library(survival)
```
A recently developed drug, MediCo, has been used over the past 12 months to treat a potentially fatal disease. 
MediCo was approved by a medical regulator last year following initial trials and mortality data has been collected over the past 12 months, since approval, to continue reviewing the drug’s effectiveness. 

This mortality data has been compared with mortality data collected from infected patients, prior to MediCo’s approval, who were NOT administered the drug. 

It is suspected that gender may be a significant covariate on the mortality rate of infected patients. 


#### Data 

The ‘CS2B_Apr_21_Qu_3_Data.csv’ file contains the combined mortality data from this investigation for 4,400 infected patients. The file contains the following five 
variables: 
* Life Unique patient identifier (integers 1, 2, …, 4,400) 
* Drug Drug indicator (1 = received drug, 0 = did not receive drug) 
* Gender Gender indicator (1 = female, 0 = male) 
* Status Status indicator (1 = death due to disease, 0 = censoring event 
occurred) 
* Time Duration in days at which death/censoring occurred (integers with a 
range of 1–365, with 1 = first day of the investigation and 
365 = last day of the investigation). 

#### Data

Before answering this question, the ‘CS2B_Apr_21_Qu_3_Data.csv’ file should be 
loaded into R and assigned to a data frame called mortalitydata. 

#### Exercises
1. Plot the Kaplan–Meier survival function estimate for all patients, together with 
its two-sided 99.5% confidence interval.  

2. Determine, using the estimated survival function from part (1. , the probability 
that a patient survived from the beginning of the investigation to the end of the 
investigation.  

3. Evaluate the appropriateness of the probability value, calculated in exercise  (2.  , 
for assessing MediCo’s effectiveness.  

4. Plot, on a single graph, four Kaplan–Meier survival function estimates without 
any confidence intervals, where each estimate represents one of the four 
possible patient group combinations of drug and gender. You should use 
separate colours to identify each survival function.  
5. Comment on the graph produced in exercise  (4. .  
<p>
6. Estimate a Cox proportional hazards model with death as the event of interest 
using the two covariates, drug and gender, with no interaction term, pasting 
your results into your answer script. You should use the Breslow method for 
tie handling.  

7. Comment on the results produced in exercise  (v1.  with reference to the effects of 
the two covariates, drug and gender, on the mortality rate. 
(v3.  Update the Cox proportional hazards model in exercise  (v1.  to include an 
interaction term between drug and gender, pasting your results into your 
answer script.  
(ix) Analyse the effectiveness of MediCo, commenting on any differences between 
males and females. [5] 
 [Total 44] 
END OF PAPER 