
%=====================================================================================================================%
\begin{frame}

For example, to load the fda package:

> library(fda)

One important thing to note is that if you terminate your session and start a new session with the saved workspace, you must load  the packages again



install.packages("evir")
 
To get out of R, just type: q(). 

\end{frame}

\begin{frame}

Section 10: Simulation
Simulation Study : Random Walks
Simulation Study: Distribution of pairwise maxima and minima
Simulation Study : Gamblers Ruins
Simulation Study: Probability of Gambler Ruin

\end{frame}
%=====================================================================================================================%
\begin{frame}

Section 1: Basic R commands and Functions
Installing R on your computer
R can be easily downloaded from the Comprehenive R Archive Network (CRAN) website.
\end{frame}
%=====================================================================================================================%
\begin{frame}

Editing your Data

x=c(0 ,5)     	      # create a vector x
data.entry(x)  	   # edit the values using spreadsheet interface.
x  	                     # print to screen
x=edit(x)	          # the 'edit' function to call the script editor
x  	                     # print to screen

\end{frame}

%=====================================================================================================================%
\begin{frame}
Manipulating Characters
> nchar("oscar")
[1] 5

Objects
During an R session, objects are created and stored by name. The command "ls()" displays all currently-stored objects (workspace). Objects can be removed using the "rm()" function.

\end{frame}
%=====================================================================================================================%
\begin{frame}

ls()
rm(x, a, temp, wt.males)
rm(list=ls())								#removes all of the objects in the workspace.


At the end of each R session, you are prompted to save your workspace. If you click Yes, all objects are written to the ".RData" file. 
When R is re-started, it reloads the workspace from this file and the command history stored in ".Rhistory" is also reloaded.

%==================================================================================================================%
