---
title: "``adorn_percentages`` Convert a data.frame of counts to percentages."
author: "Kevin O'Brien"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(janitor)
```


#### Description
This function excludes the first column of the input data.frame, assuming that it contains a descriptive
variable. If the input data.frame is not a tabyl, it will convert to one in order to preserve the
underlying values in the core attribute.

#### Usage 
<pre><code>
adorn_percentages(dat, denominator = "row", na.rm = TRUE)
</code></pre> 

#### Arguments
* ``dat``: a tabyl or other data.frame with a tabyl-like layout. If given a list of data.frames,
this function will apply itself to each data.frame in the list (designed for 3-way tabyl lists).
* ``denominator``: the direction to use for calculating percentages. One of "row", "col", or "all".
* ``na.rm``: should missing values (including NaN) be omitted from the calculations?

#### Value
Returns a data.frame of percentages, expressed as numeric values between 0 and 1.

#### Examples
```{r}
mtcars %>%
tabyl(am, cyl) %>%
adorn_percentages("col")
# calculates correctly even with totals column and/or row:
```
```{r}
mtcars %>%
tabyl(am, cyl) %>%
adorn_totals("row") %>%
adorn_percentages()
```
