---
title: "janitor"
author: "Kevin O'Brien"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(janitor)
```

18 tabyl
tabyl Generate a frequency table (1-, 2-, or 3-way).
#### Description
A fully-featured alternative to table(). Results are data.frames and can be formatted and enhanced
with janitor’s family of adorn_ functions.
Specify a data.frame and the one, two, or three unquoted column names you want to tabulate. Three
variables generates a list of 2-way tabyls, split by the third variable.
Alternatively, you can tabulate a single variable that isn’t in a data.frame by calling tabyl on a
vector, e.g., tabyl(mtcars$gear).
#### Usage 
<pre><code>
tabyl(dat, ...)
## Default S3 method:
tabyl(dat, show_na = TRUE,
show_missing_levels = TRUE, ...)
## S3 method for class 'data.frame'
tabyl(dat, var1, var2, var3, show_na = TRUE,
show_missing_levels = TRUE, ...)
</code></pre> 

#### Arguments

* ``dat``: a data.frame containing the variables you wish to count. Or, a vector you want
to tabulate.
... the </code></pre> #### Arguments to tabyl (here just for the sake of documentation compliance, as
all </code></pre> #### Arguments are listed with the vector- and data.frame-specific methods)
show_na should counts of NA values be displayed? In a one-way tabyl, the presence of
NA values triggers an additional column showing valid percentages(calculated
excluding NA values).
show_missing_levels
should counts of missing levels of factors be displayed? These will be rows
and/or columns of zeroes. Useful for keeping consistent output dimensions even
when certain factor levels may not be present in the data.
var1 the column name of the first variable.
var2 (optional) the column name of the second variable (the rows in a 2-way tabulation).
var3 (optional) the column name of the third variable (the list in a 3-way tabulation).


#### Examples
```{r}
tabyl(mtcars, cyl)
tabyl(mtcars, cyl, gear)
tabyl(mtcars, cyl, gear, am)
# or using the %>% pipe
mtcars %>%
tabyl(cyl, gear)
# illustrating show_na functionality:
my_cars <- rbind(mtcars, rep(NA, 11))
my_cars %>% tabyl(cyl)
my_cars %>% tabyl(cyl, show_na = FALSE)
# Calling on a single vector not in a data.frame:
val <- c("hi", "med", "med", "lo")
tabyl(val)
```

#### Value
Returns a data.frame with frequencies and percentages of the tabulated variable(s). A 3-way tabulation
returns a list of data.frames.
top_levels 19
